---
title: "simstudy_PART5_fs_dp_compare"
author: "Kelsey"
date: "10/5/2021"
output: html_document
---

```{r, }
require(knitr)
```
```{r}
#get the needed packages
if(!require("pacman")){
	install.packages("pacman")
	library(pacman)}
p_load("dplyr", "ggplot2", "viridis", "tidyr",  "fGarch", "sn", "purrr", "forcats", "stats", "changepoint", "multipanelfigure")
```
```{r}
sim_spp <- read.csv(file = "sim_spp.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F)
accumulated <- read.csv(file = "accumulated.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F)

accumulated
```

```{r}
#accumulated$Taxa_Naturalized_centered <- accumulated$Taxa_Naturalized - mean(accumulated$Taxa_Naturalized)
plot(Taxa_Naturalized ~  Year, data = accumulated)
```
```{r}
#accumulated$Taxa_Naturalized_centered <- accumulated$Taxa_Naturalized - mean(accumulated$Taxa_Naturalized)
hist(sim_spp$Naturalization_Year)
```

### Changepoint Analysis

```{r}
a2 <- changepoint::cpt.meanvar(accumulated$Taxa_Naturalized, penalty = "MBIC", method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)
```

```{r}
changepoint::plot(a2, type = "p")
```
```{r}
changepoint::param.est(a2)
```
```{r}
accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
```

# What if no lags exist??


<!-- ## Simulate Hypothetical Lag Times -->

<!-- For this scenario, let's say that all lags have a mean of 10 years -->

<!-- ```{r} -->
<!-- sim_spp12 <- sim_spp -->
<!-- ``` -->

<!-- ##### 3.0 -->

<!-- Simulate FIELD SURVEYING lags -->

<!-- ```{r} -->
<!-- mean = 0 -->
<!-- SD = 0 -->
<!-- skew = 0 -->
<!-- params <- sn::cp2dp(c(mean, SD, skew), "SN") -->

<!-- FS_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0)) -->
<!-- colnames(FS_Lag) <- "FS_Lag" -->
<!-- #FS_Lag <- subset(FS_Lag, FS_Lag >=0) -->
<!-- FS_Lag <- as.data.frame(sample(FS_Lag$FS_Lag, 1000)) -->
<!-- colnames(FS_Lag) <- "FS_Lag" -->

<!-- #plotting histogram -->
<!--  ggplot(FS_Lag, aes(x = FS_Lag)) + -->
<!--   geom_histogram(binwidth = 1) + -->
<!--    theme_classic() -->
<!--  summary(FS_Lag) -->
<!-- FS_Lag -->
<!-- ``` -->

<!-- ##### 3.1 -->
<!-- Simulate DATA PROCESSING lags -->

<!-- ```{r} -->
<!-- mean = 0 -->
<!-- SD = 0 -->
<!-- skew = 0 -->
<!-- params <- sn::cp2dp(c(mean, SD, skew), "SN") -->

<!-- DP_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0)) -->
<!-- colnames(DP_Lag) <- "DP_Lag" -->
<!-- #FS_Lag <- subset(FS_Lag, FS_Lag >=0) -->
<!-- DP_Lag <- as.data.frame(sample(DP_Lag$DP_Lag, 1000)) -->
<!-- colnames(DP_Lag) <- "DP_Lag" -->

<!-- #plotting histogram -->
<!--  ggplot(DP_Lag, aes(x = DP_Lag)) + -->
<!--   geom_histogram(binwidth = 1) + -->
<!--    theme_classic() -->
<!--  summary(DP_Lag) -->
<!-- DP_Lag -->
<!-- ``` -->


<!-- ##### 3.6 Okay, lets add up the lags -->
<!-- So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag, -->

<!-- All created by adding or subtracting from the naturalization year -->
<!-- ```{r} -->
<!-- sim_spp12$Voucher_Year <- sim_spp12$Naturalization_Year + FS_Lag$FS_Lag -->
<!-- sim_spp12$Report_Year <- sim_spp12$Naturalization_Year + FS_Lag$FS_Lag + DP_Lag$DP_Lag -->

<!-- sim_spp12 <- sim_spp12 %>% dplyr::select(Species_IDNo,  Naturalization_Year, Voucher_Year,  Report_Year) -->
<!-- sim_spp12 -->
<!-- ``` -->

<!-- ### Cropping the Dataset because we can't know about species that haven't been reported yet -->

<!-- What's the lags look like for all species -->
<!-- ```{r} -->
<!-- # #plotting histogram -->
<!-- # ggplot(sim_spp12, aes(x = All_Processing_Lags)) + -->
<!-- #   geom_density(fill="#440154FF", color="#e9ecef", alpha=0.6) + -->
<!-- #  # scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) + -->
<!-- #   theme_classic() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- sim_spp12_crop <- subset(sim_spp12, Report_Year < 2020) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- nrow(sim_spp12) -nrow(sim_spp12_crop) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- retrieval_rate = 1 -->
<!-- sampled <- as.data.frame(sample(sim_spp12_crop$Species_IDNo, (nrow(sim_spp12_crop)* retrieval_rate))) -->
<!-- colnames(sampled) <- "Species_IDNo" -->
<!-- sim_spp12_crop_sample <- merge(sampled, sim_spp12_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE) -->
<!-- sim_spp12_crop_sample -->
<!-- ``` -->


<!-- ## Step 4. Analyse Trends - A) Voucher Year -->
<!-- ```{r} -->
<!-- #just making a vector of years from 1920 to 2020 -->
<!-- alltheyears <- as.data.frame(seq(from = 1920, to = 2019)) -->
<!-- colnames(alltheyears) <- "Year" -->

<!-- #grouping my simulated species by the year they we vouchered -->
<!-- year_count <- as.data.frame(table(sim_spp12_crop_sample$Voucher_Year)) %>% -->
<!-- # and renaming columns -->
<!--   dplyr::rename(Year = Var1, Taxa_Vouchered = Freq) -->

<!-- #merging with all the years -->
<!-- year_count <-merge(alltheyears, year_count, all = TRUE) -->
<!-- #making the NA's become 0s, because 0 new species were vouchered in those years -->
<!-- year_count[is.na(year_count)] = 0 -->

<!-- #calculating species accumulation over time -->
<!-- accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum) -->
<!-- accumulated_A <- cbind(year_count, accumulated_A) -->
<!-- accumulated_A -->
<!-- ``` -->


<!-- #### Changepoint - NO LAGS AT ALL -->

<!-- ```{r} -->
<!-- a2 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "MBIC", method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- png("nolags_dotplot.png", width = 154, height = 114, units='mm', res = 600) -->
<!-- changepoint::plot(a2, type = "p", ylim = c(0,20), pch = 1, cex = 1.25, cpt.width = 3, col = "red", cpt.col = "black", ylab = "Naturalization Rate (# species/year) ", xlab = "Time (Years Since Beginning of Analysis Window)") -->
<!-- text(x=46, y=10.6, "10 +/- 3.5", cex = 1, font = 2) -->
<!-- dev.off() -->
<!-- ``` -->
<!-- ```{r} -->
<!-- changepoint::param.est(a2)$mean -->
<!-- sqrt(changepoint::param.est(a2)$variance) -->
<!-- ``` -->



<!-- ```{r} -->
<!-- #making into a single dataframe -->
<!-- #accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized) -->
<!-- accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered) -->
<!-- accumulated$Analysis_Type <- "Hypothetical Naturalization Date" -->
<!-- accumulated_A$Analysis_Type <- "Evidence Collection Date" -->

<!-- accumulated_A_B_C_D <- rbind(accumulated, accumulated_A) -->
<!-- accumulated_A_B_C_D -->
<!-- ``` -->

<!-- ```{r} -->
<!-- DATA <- accumulated_A_B_C_D# %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species)) -->
<!-- DATA <- subset(DATA, DATA$Analysis_Type != "Hypothetical Naturalization Date") -->
<!-- #DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date") -->
<!-- #plotting total species accumulation over time -->
<!-- ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) + -->
<!--   geom_line( color = "black", size=2, alpha=1) + -->
<!--  # scale_color_viridis(discrete = TRUE) + -->
<!--   scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) + -->
<!--   guides(colour = guide_legend(reverse=T)) + -->
<!--  theme_classic() + -->
<!--   theme(legend.title=element_blank()) + -->
<!--   labs( title = "A:  No Lags Exist", x = "Year", y = "Total Naturalized Species") -->
<!-- ggsave(filename = "nolags_curve.png", device = "png", scale = 1, width =6, dpi = 600) -->
<!-- ``` -->







# What if only field Surveying Lags Exist??


<!-- ## Simulate Hypothetical Lag Times -->

<!-- For this scenario, let's say that all lags have a mean of 10 years -->

<!-- ```{r} -->
<!-- sim_spp13 <- sim_spp -->
<!-- ``` -->

<!-- ##### 3.0 -->

<!-- Simulate FIELD SURVEYING lags -->

<!-- ```{r} -->
<!-- mean = 10 -->
<!-- SD = 3.5 -->
<!-- skew = 0 -->
<!-- params <- sn::cp2dp(c(mean, SD, skew), "SN") -->

<!-- FS_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0)) -->
<!-- colnames(FS_Lag) <- "FS_Lag" -->
<!-- #FS_Lag <- subset(FS_Lag, FS_Lag >=0) -->
<!-- FS_Lag <- as.data.frame(sample(FS_Lag$FS_Lag, 1000)) -->
<!-- colnames(FS_Lag) <- "FS_Lag" -->

<!-- #plotting histogram -->
<!--  ggplot(FS_Lag, aes(x = FS_Lag)) + -->
<!--   geom_histogram(binwidth = 1) + -->
<!--    theme_classic() -->
<!--  summary(FS_Lag) -->
<!-- FS_Lag -->
<!-- ``` -->

<!-- ##### 3.1 -->
<!-- Simulate DATA PROCESSING lags -->

<!-- ```{r} -->
<!-- mean = 0 -->
<!-- SD = 0 -->
<!-- skew = 0 -->
<!-- params <- sn::cp2dp(c(mean, SD, skew), "SN") -->

<!-- DP_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0)) -->
<!-- colnames(DP_Lag) <- "DP_Lag" -->
<!-- #FS_Lag <- subset(FS_Lag, FS_Lag >=0) -->
<!-- DP_Lag <- as.data.frame(sample(DP_Lag$DP_Lag, 1000)) -->
<!-- colnames(DP_Lag) <- "DP_Lag" -->

<!-- #plotting histogram -->
<!--  ggplot(DP_Lag, aes(x = DP_Lag)) + -->
<!--   geom_histogram(binwidth = 1) + -->
<!--    theme_classic() -->
<!--  summary(DP_Lag) -->
<!-- DP_Lag -->
<!-- ``` -->


<!-- ##### 3.6 Okay, lets add up the lags -->
<!-- So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag, -->

<!-- All created by adding or subtracting from the naturalization year -->
<!-- ```{r} -->
<!-- sim_spp13$Voucher_Year <- sim_spp13$Naturalization_Year + FS_Lag$FS_Lag -->
<!-- sim_spp13$Report_Year <- sim_spp13$Naturalization_Year + FS_Lag$FS_Lag + DP_Lag$DP_Lag -->

<!-- sim_spp13 <- sim_spp13 %>% dplyr::select(Species_IDNo,  Naturalization_Year, Voucher_Year,  Report_Year) -->
<!-- sim_spp13 -->
<!-- ``` -->

<!-- ### Cropping the Dataset because we can't know about species that haven't been reported yet -->

<!-- What's the lags look like for all species -->
<!-- ```{r} -->
<!-- # #plotting histogram -->
<!-- # ggplot(sim_spp13, aes(x = All_Processing_Lags)) + -->
<!-- #   geom_density(fill="#440154FF", color="#e9ecef", alpha=0.6) + -->
<!-- #  # scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) + -->
<!-- #   theme_classic() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- sim_spp13_crop <- subset(sim_spp13, Report_Year < 2020) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- nrow(sim_spp13) -nrow(sim_spp13_crop) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- retrieval_rate = 1 -->
<!-- sampled <- as.data.frame(sample(sim_spp13_crop$Species_IDNo, (nrow(sim_spp13_crop)* retrieval_rate))) -->
<!-- colnames(sampled) <- "Species_IDNo" -->
<!-- sim_spp13_crop_sample <- merge(sampled, sim_spp13_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE) -->
<!-- sim_spp13_crop_sample -->
<!-- ``` -->


<!-- ## Step 4. Analyse Trends - A) Voucher Year -->
<!-- ```{r} -->
<!-- #just making a vector of years from 1920 to 2020 -->
<!-- alltheyears <- as.data.frame(seq(from = 1920, to = 2019)) -->
<!-- colnames(alltheyears) <- "Year" -->

<!-- #grouping my simulated species by the year they we vouchered -->
<!-- year_count <- as.data.frame(table(sim_spp13_crop_sample$Voucher_Year)) %>% -->
<!-- # and renaming columns -->
<!--   dplyr::rename(Year = Var1, Taxa_Vouchered = Freq) -->

<!-- #merging with all the years -->
<!-- year_count <-merge(alltheyears, year_count, all = TRUE) -->
<!-- #making the NA's become 0s, because 0 new species were vouchered in those years -->
<!-- year_count[is.na(year_count)] = 0 -->

<!-- #calculating species accumulation over time -->
<!-- accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum) -->
<!-- accumulated_A <- cbind(year_count, accumulated_A) -->
<!-- accumulated_A -->
<!-- ``` -->



<!-- ### Changpoints -->


<!-- ```{r} -->
<!-- plot(accumulated_A$Taxa_Vouchered) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- a1 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- plot(a1, diagnostic = TRUE) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- selectedcp = 2 -->
<!-- #changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)") -->
<!-- changepoint::plot(a1, ncpts = selectedcp) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- changepoint::cpts.full(a1) + (accumulated_A$Year[1] -1) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- firstrow = 3 -->
<!-- secondrow = firstrow +1 -->
<!-- a2 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "Manual",  pen.value=((pen.value.full(a1)[firstrow] + pen.value.full(a1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A)) -->
<!-- chpoints -->
<!-- chpoints + accumulated_A$Year[1] -->
<!-- segmeans <- changepoint::param.est(a2)$mean -->
<!-- segmeans -->
<!-- segvars <- (changepoint::param.est(a2)$variance) -->
<!-- sqrt(segvars) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8]) -->
<!-- seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8]) -->

<!-- segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean) -->
<!-- segwmeans <- segwmeans[!is.na(segwmeans)] -->
<!-- segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar) -->
<!-- segwvars <- segwvars[!is.na(segwvars)] -->
<!-- ``` -->
<!-- ```{r} -->
<!-- startptindex = 3 -->
<!-- endptindex = 4 -->

<!-- print("How many species did we lose that hadn't been reported yet?") -->
<!-- spplostharvest <- nrow(sim_spp13) -nrow(sim_spp13_crop) -->
<!-- spplostharvest -->

<!-- spplostretrieval <- nrow(sim_spp13_crop) - nrow(sim_spp13_crop_sample) -->
<!-- spplostretrieval -->

<!-- print("start seg length, mean, std: ") -->
<!-- chpoints[startptindex] -->
<!-- round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2) -->
<!-- round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2) -->

<!-- print("truncated seg length, mean, std:") -->
<!-- chpoints[endptindex] - chpoints[startptindex] -->
<!-- round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2) -->

<!-- print("ending seg length, mean, std:") -->
<!-- nrow(accumulated_A) - chpoints[endptindex] -->
<!-- round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A)) -->
<!-- chpoints -->
<!-- chpoints + accumulated_A$Year[1] -->
<!-- segmeans <- changepoint::param.est(a2)$mean -->
<!-- segmeans -->
<!-- segvars <- (changepoint::param.est(a2)$variance) -->
<!-- sqrt(segvars) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- mean(accumulated_A$Taxa_Vouchered) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- png("FSlags_dotplot.png", width = 154, height = 114, units='mm', res = 600) -->
<!-- changepoint::plot(a2, ylim = c(0,20), type = "p", pch = 1, cex = 1.25, cpt.width = 3, col = "red", cpt.col = "black", ylab = "Naturalization Rate (# species/year) ", xlab = "Time (Years Since Beginning of Analysis Window)") -->
<!-- text(x=16, y=0.9, "0.29 +/- 0.45", , cex = 1, font = 2) -->
<!-- text(x=21, y= 4.2, "3.80 +/- 1.60", , cex = 1, font = 2) -->
<!-- #text(x=40, y=5.6,  "4.90 +/- 2.47", cex = 1, font = 2) -->
<!-- text(x=62, y=10.8, "10.07 +/- 3.29", cex = 1, font = 2) -->
<!-- dev.off() -->
<!-- ``` -->


<!-- ```{r} -->
<!-- #making into a single dataframe -->
<!-- #accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized) -->
<!-- accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered) -->
<!-- accumulated$Analysis_Type <- "Hypothetical Naturalization Date" -->
<!-- accumulated_A$Analysis_Type <- "Evidence Collection Date" -->

<!-- accumulated_A_B_C_D <- rbind(accumulated, accumulated_A) -->
<!-- accumulated_A_B_C_D -->
<!-- ``` -->
<!-- ```{r} -->
<!-- DATA <- accumulated_A_B_C_D# %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species)) -->
<!-- DATA <- subset(DATA, DATA$Analysis_Type != "Hypothetical Naturalization Date") -->
<!-- #DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date") -->
<!-- #plotting total species accumulation over time -->
<!-- ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) + -->
<!--   geom_line( color = "black", size=2, alpha=1) + -->
<!--  # scale_color_viridis(discrete = TRUE) + -->
<!--   scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) + -->
<!--   guides(colour = guide_legend(reverse=T)) + -->
<!--  theme_classic() + -->
<!--   theme(legend.title=element_blank()) + -->
<!--   labs( title = "B:  Only Field Surveying Lags Exist", x = "Year", y = "Total Naturalized Species") -->
<!-- ggsave(filename = "FSlags_curve.png", device = "png", scale = 1, width =6, dpi = 600) -->
<!-- ``` -->




# What if only Data Processing Lags Exist??


<!-- ## Simulate Hypothetical Lag Times -->

<!-- For this scenario, let's say that all lags have a mean of 10 years -->

<!-- ```{r} -->
<!-- sim_spp14 <- sim_spp -->
<!-- ``` -->

<!-- ##### 3.0 -->

<!-- Simulate FIELD SURVEYING lags -->

<!-- ```{r} -->
<!-- mean = 0 -->
<!-- SD = 0 -->
<!-- skew = 0 -->
<!-- params <- sn::cp2dp(c(mean, SD, skew), "SN") -->

<!-- FS_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0)) -->
<!-- colnames(FS_Lag) <- "FS_Lag" -->
<!-- #FS_Lag <- subset(FS_Lag, FS_Lag >=0) -->
<!-- FS_Lag <- as.data.frame(sample(FS_Lag$FS_Lag, 1000)) -->
<!-- colnames(FS_Lag) <- "FS_Lag" -->

<!-- #plotting histogram -->
<!--  ggplot(FS_Lag, aes(x = FS_Lag)) + -->
<!--   geom_histogram(binwidth = 1) + -->
<!--    theme_classic() -->
<!--  summary(FS_Lag) -->
<!-- FS_Lag -->
<!-- ``` -->

<!-- ##### 3.1 -->
<!-- Simulate DATA PROCESSING lags -->

<!-- ```{r} -->
<!-- mean = 10 -->
<!-- SD = 3.5 -->
<!-- skew = 0 -->
<!-- params <- sn::cp2dp(c(mean, SD, skew), "SN") -->

<!-- DP_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0)) -->
<!-- colnames(DP_Lag) <- "DP_Lag" -->
<!-- #FS_Lag <- subset(FS_Lag, FS_Lag >=0) -->
<!-- DP_Lag <- as.data.frame(sample(DP_Lag$DP_Lag, 1000)) -->
<!-- colnames(DP_Lag) <- "DP_Lag" -->

<!-- #plotting histogram -->
<!--  ggplot(DP_Lag, aes(x = DP_Lag)) + -->
<!--   geom_histogram(binwidth = 1) + -->
<!--    theme_classic() -->
<!--  summary(DP_Lag) -->
<!-- DP_Lag -->
<!-- ``` -->


<!-- ##### 3.6 Okay, lets add up the lags -->
<!-- So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag, -->

<!-- All created by adding or subtracting from the naturalization year -->
<!-- ```{r} -->
<!-- sim_spp14$Voucher_Year <- sim_spp14$Naturalization_Year + FS_Lag$FS_Lag -->
<!-- sim_spp14$Report_Year <- sim_spp14$Naturalization_Year + FS_Lag$FS_Lag + DP_Lag$DP_Lag -->

<!-- sim_spp14 <- sim_spp14 %>% dplyr::select(Species_IDNo,  Naturalization_Year, Voucher_Year,  Report_Year) -->
<!-- sim_spp14 -->
<!-- ``` -->

<!-- ### Cropping the Dataset because we can't know about species that haven't been reported yet -->

<!-- What's the lags look like for all species -->
<!-- ```{r} -->
<!-- # #plotting histogram -->
<!-- # ggplot(sim_spp14, aes(x = All_Processing_Lags)) + -->
<!-- #   geom_density(fill="#440154FF", color="#e9ecef", alpha=0.6) + -->
<!-- #  # scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) + -->
<!-- #   theme_classic() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- sim_spp14_crop <- subset(sim_spp14, Report_Year < 2020) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- nrow(sim_spp14) -nrow(sim_spp14_crop) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- retrieval_rate = 1 -->
<!-- sampled <- as.data.frame(sample(sim_spp14_crop$Species_IDNo, (nrow(sim_spp14_crop)* retrieval_rate))) -->
<!-- colnames(sampled) <- "Species_IDNo" -->
<!-- sim_spp14_crop_sample <- merge(sampled, sim_spp14_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE) -->
<!-- sim_spp14_crop_sample -->
<!-- ``` -->


<!-- ## Step 4. Analyse Trends - A) Voucher Year -->
<!-- ```{r} -->
<!-- #just making a vector of years from 1920 to 2020 -->
<!-- alltheyears <- as.data.frame(seq(from = 1920, to = 2019)) -->
<!-- colnames(alltheyears) <- "Year" -->

<!-- #grouping my simulated species by the year they we vouchered -->
<!-- year_count <- as.data.frame(table(sim_spp14_crop_sample$Voucher_Year)) %>% -->
<!-- # and renaming columns -->
<!--   dplyr::rename(Year = Var1, Taxa_Vouchered = Freq) -->

<!-- #merging with all the years -->
<!-- year_count <-merge(alltheyears, year_count, all = TRUE) -->
<!-- #making the NA's become 0s, because 0 new species were vouchered in those years -->
<!-- year_count[is.na(year_count)] = 0 -->

<!-- #calculating species accumulation over time -->
<!-- accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum) -->
<!-- accumulated_A <- cbind(year_count, accumulated_A) -->
<!-- accumulated_A -->
<!-- ``` -->



<!-- ### Changpoints -->


<!-- ```{r} -->
<!-- plot(accumulated_A$Taxa_Vouchered) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- a1 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- plot(a1, diagnostic = TRUE) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- selectedcp = 2 -->
<!-- #changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)") -->
<!-- changepoint::plot(a1, ncpts = selectedcp) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- changepoint::cpts.full(a1) + (accumulated_A$Year[1] -1) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- firstrow =  -->
<!-- secondrow = firstrow +1 -->
<!-- a2 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "Manual",  pen.value=((pen.value.full(a1)[firstrow] + pen.value.full(a1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A)) -->
<!-- chpoints -->
<!-- chpoints + accumulated_A$Year[1] -->
<!-- segmeans <- changepoint::param.est(a2)$mean -->
<!-- segmeans -->
<!-- segvars <- (changepoint::param.est(a2)$variance) -->
<!-- sqrt(segvars) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8]) -->
<!-- seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8]) -->

<!-- segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean) -->
<!-- segwmeans <- segwmeans[!is.na(segwmeans)] -->
<!-- segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar) -->
<!-- segwvars <- segwvars[!is.na(segwvars)] -->
<!-- ``` -->
<!-- ```{r} -->
<!-- startptindex = 1 -->
<!-- endptindex = 2 -->

<!-- print("How many species did we lose that hadn't been reported yet?") -->
<!-- spplostharvest <- nrow(sim_spp14) -nrow(sim_spp14_crop) -->
<!-- spplostharvest -->

<!-- spplostretrieval <- nrow(sim_spp14_crop) - nrow(sim_spp14_crop_sample) -->
<!-- spplostretrieval -->

<!-- print("start seg length, mean, std: ") -->
<!-- chpoints[startptindex] -->
<!-- round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2) -->
<!-- round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2) -->

<!-- print("truncated seg length, mean, std:") -->
<!-- chpoints[endptindex] - chpoints[startptindex] -->
<!-- round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2) -->

<!-- print("ending seg length, mean, std:") -->
<!-- nrow(accumulated_A) - chpoints[endptindex] -->
<!-- round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A)) -->
<!-- chpoints -->
<!-- chpoints + accumulated_A$Year[1] -->
<!-- segmeans <- changepoint::param.est(a2)$mean -->
<!-- segmeans -->
<!-- segvars <- (changepoint::param.est(a2)$variance) -->
<!-- sqrt(segvars) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- mean(accumulated_A$Taxa_Vouchered) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- png("DPlags_dotplot.png", width = 154, height = 114, units='mm', res = 600) -->
<!-- changepoint::plot(a2, ylim = c(0,20), type = "p", pch = 1, cex = 1.25, cpt.width = 3, col = "red", cpt.col = "black", ylab = "Naturalization Rate (# species/year) ", xlab = "Time (Years Since Beginning of Analysis Window)") -->
<!-- text(x=83, y=0.6, "0.33 +/- 0.47", , cex = 1, font = 2) -->
<!-- text(x=93, y= 5.6, "4.83 +/- 2.03", , cex = 1, font = 2) -->
<!-- #text(x=40, y=5.6,  "4.90 +/- 2.47", cex = 1, font = 2) -->
<!-- text(x=48, y=10.7, "10.01 +/- 3.54", cex = 1, font = 2) -->
<!-- dev.off() -->
<!-- ``` -->


<!-- ```{r} -->
<!-- #making into a single dataframe -->
<!-- #accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized) -->
<!-- accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered) -->
<!-- accumulated$Analysis_Type <- "Hypothetical Naturalization Date" -->
<!-- accumulated_A$Analysis_Type <- "Evidence Collection Date" -->

<!-- accumulated_A_B_C_D <- rbind(accumulated, accumulated_A) -->
<!-- accumulated_A_B_C_D -->
<!-- ``` -->
<!-- ```{r} -->
<!-- DATA <- accumulated_A_B_C_D# %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species)) -->
<!-- DATA <- subset(DATA, DATA$Analysis_Type != "Hypothetical Naturalization Date") -->
<!-- #DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date") -->
<!-- #plotting total species accumulation over time -->
<!-- ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) + -->
<!--   geom_line( color = "black", size=2, alpha=1) + -->
<!--  # scale_color_viridis(discrete = TRUE) + -->
<!--   scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) + -->
<!--   guides(colour = guide_legend(reverse=T)) + -->
<!--  theme_classic() + -->
<!--   theme(legend.title=element_blank()) + -->
<!--   labs( title = "C:  Only Data Processing Lags Exist", x = "Year", y = "Total Naturalized Species") -->
<!-- ggsave(filename = "DPlags_curve.png", device = "png", scale = 1, width =6, dpi = 600) -->
<!-- ``` -->




# What if BOTH Field Surveying and Data Processing Lags Exist?


## Simulate Hypothetical Lag Times

For this scenario, let's say that all lags have a mean of 10 years

```{r}
sim_spp15 <- sim_spp
```

##### 3.0

Simulate FIELD SURVEYING lags

```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

FS_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(FS_Lag) <- "FS_Lag"
#FS_Lag <- subset(FS_Lag, FS_Lag >=0)
FS_Lag <- as.data.frame(sample(FS_Lag$FS_Lag, 1000))
colnames(FS_Lag) <- "FS_Lag"

#plotting histogram
 ggplot(FS_Lag, aes(x = FS_Lag)) +
  geom_histogram(binwidth = 1) +
   theme_classic()
 summary(FS_Lag)
FS_Lag
```

##### 3.1
Simulate DATA PROCESSING lags

```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

DP_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(DP_Lag) <- "DP_Lag"
#FS_Lag <- subset(FS_Lag, FS_Lag >=0)
DP_Lag <- as.data.frame(sample(DP_Lag$DP_Lag, 1000))
colnames(DP_Lag) <- "DP_Lag"

#plotting histogram
 ggplot(DP_Lag, aes(x = DP_Lag)) +
  geom_histogram(binwidth = 1) +
   theme_classic()
 summary(DP_Lag)
DP_Lag
```


##### 3.6 Okay, lets add up the lags
So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag,

All created by adding or subtracting from the naturalization year
```{r}
sim_spp15$Voucher_Year <- sim_spp15$Naturalization_Year + FS_Lag$FS_Lag
sim_spp15$Report_Year <- sim_spp15$Naturalization_Year + FS_Lag$FS_Lag + DP_Lag$DP_Lag

sim_spp15 <- sim_spp15 %>% dplyr::select(Species_IDNo,  Naturalization_Year, Voucher_Year,  Report_Year)
sim_spp15
```

### Cropping the Dataset because we can't know about species that haven't been reported yet

What's the lags look like for all species
```{r}
# #plotting histogram
# ggplot(sim_spp15, aes(x = All_Processing_Lags)) +
#   geom_density(fill="#440154FF", color="#e9ecef", alpha=0.6) +
#  # scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
#   theme_classic()
```

```{r}
sim_spp15_crop <- subset(sim_spp15, Report_Year < 2020)
```

```{r}
nrow(sim_spp15) -nrow(sim_spp15_crop)
```
```{r}
retrieval_rate = 1
sampled <- as.data.frame(sample(sim_spp15_crop$Species_IDNo, (nrow(sim_spp15_crop)* retrieval_rate)))
colnames(sampled) <- "Species_IDNo"
sim_spp15_crop_sample <- merge(sampled, sim_spp15_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE)
sim_spp15_crop_sample
```


## Step 4. Analyse Trends - A) Voucher Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(sim_spp15_crop_sample$Voucher_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Vouchered = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum)
accumulated_A <- cbind(year_count, accumulated_A)
accumulated_A
```



### Changpoints


```{r}
plot(accumulated_A$Taxa_Vouchered)
```
```{r}
a1 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

plot(a1, diagnostic = TRUE)
```
```{r}
selectedcp = 4
#changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)")
changepoint::plot(a1, ncpts = selectedcp)
```
```{r}
changepoint::cpts.full(a1) + (accumulated_A$Year[1] -1)
```
```{r}
firstrow = 5
secondrow = firstrow +1
a2 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "Manual",  pen.value=((pen.value.full(a1)[firstrow] + pen.value.full(a1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A))
chpoints
chpoints + accumulated_A$Year[1]
segmeans <- changepoint::param.est(a2)$mean
segmeans
segvars <- (changepoint::param.est(a2)$variance)
sqrt(segvars)
```

```{r}
seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1])
seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2])
seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3])
seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4])
seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5])
seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6])
seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7])
seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8])
seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1])
seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2])
seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3])
seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4])
seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5])
seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6])
seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7])
seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8])

segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean)
segwmeans <- segwmeans[!is.na(segwmeans)]
segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar)
segwvars <- segwvars[!is.na(segwvars)]
```
```{r}
startptindex = 3
endptindex = 4

print("How many species did we lose that hadn't been reported yet?")
spplostharvest <- nrow(sim_spp15) -nrow(sim_spp15_crop)
spplostharvest

spplostretrieval <- nrow(sim_spp15_crop) - nrow(sim_spp15_crop_sample)
spplostretrieval

print("start seg length, mean, std: ")
chpoints[startptindex]
round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2)
round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2)

print("truncated seg length, mean, std:")
chpoints[endptindex] - chpoints[startptindex]
round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2)
round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2)

print("ending seg length, mean, std:")
nrow(accumulated_A) - chpoints[endptindex]
round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2)
round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2)
```
```{r}
mean(accumulated_A$Taxa_Vouchered)
```

```{r}
chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A))
chpoints
chpoints + accumulated_A$Year[1]
segmeans <- changepoint::param.est(a2)$mean
segmeans
segvars <- (changepoint::param.est(a2)$variance)
sqrt(segvars)
```
```{r}
png("bothlags_dotplot.png", width = 154, height = 114, units='mm', res = 600)
changepoint::plot(a2, ylim = c(0,20), type = "p", pch = 1, cex = 1.25, cpt.width = 3, col = "red", cpt.col = "black", ylab = "Naturalization Rate (# species/year) ", xlab = "Time (Years Since Beginning of Analysis Window)")
text(x=19, y=0.5, "0.37 +/- 0.48", , cex = 1, font = 2)
text(x=09, y= 6.1, "6.72 +/- 3.22", , cex = 1, font = 2)
#text(x=40, y=5.6,  "4.90 +/- 2.47", cex = 1, font = 2)
text(x=45, y=11.2, "10.46 +/- 3.27", cex = 1, font = 2)

text(x=82, y= 5.2, "5.90 +/- 2.50", , cex = 1, font = 2)
text(x=83, y=0.5, "0.33 +/- 0.47", , cex = 1, font = 2)
dev.off()
```


```{r}
#making into a single dataframe
#accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered)
accumulated$Analysis_Type <- "Hypothetical Naturalization Date"
accumulated_A$Analysis_Type <- "Evidence Collection Date"

accumulated_A_B_C_D <- rbind(accumulated, accumulated_A)
accumulated_A_B_C_D
```
```{r}
DATA <- accumulated_A_B_C_D# %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
DATA <- subset(DATA, DATA$Analysis_Type != "Hypothetical Naturalization Date")
#DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) +
  geom_line( color = "black", size=2, alpha=1) +
 # scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
  guides(colour = guide_legend(reverse=T)) +
 theme_classic() +
  theme(legend.title=element_blank()) +
  labs( title = "D:  Both Field Surveying & Data Processing Lags Exist", x = "Year", y = "Total Naturalized Species")
ggsave(filename = "bothlags_curve.png", device = "png", scale = 1, width =6, dpi = 600)
```
