---
title: "Simstudy-PART 1-Hypothetical_Species_Creation"
author: "Kelsey"
date: "9/12/2021"
output: html_document
---

```{r, }
require(knitr)
```
```{r}
#knitr::opts_knit$set(root.dir = "C:/Users/Kelsey/Documents/codingwork/simulation_study") #set the working directory
```
```{r}
#get the needed packages
if(!require("pacman")){
	install.packages("pacman")
	library(pacman)}
p_load("dplyr", "ggplot2", "viridis", "tidyr",  "fGarch", "sn", "purrr", "forcats", "stats", "changepoint", "changepoint.np")
```


### Hypothetical Reality = Species are naturalizing at a  linear rate
#### Step 1: Build Hypothetical Reality - Creating Species

Here, I'm going to create 1000 hypothetical species that naturalized between 1920 and 2020 using a random number generator where each year has an equal chance of coming up. 

```{r}
 #creating the naturalized species, rounding to the nearest whole number (eg. year 1999 instead of 1999.433)
 sim_spp <- round(runif(1000, min = 1920, max = 2019), digits = 0)
 # and sort earliest to latest
 sim_spp <- as.data.frame(sort(sim_spp, decreasing = FALSE))
 #renaming the columns
 colnames(sim_spp) <- "Naturalization_Year"
 #naming the species a number (1 through 1000)
 sim_spp$Species_IDNo <- seq(from = 1, to = nrow(sim_spp))
```

 This histogram represents a scenario where the number of species varies per year, but there's no trend (increasing or decreasing) in the number naturalizing over time.

 #### Step 2: Simulate Hypothetical Reality - Creating Species Accumulation Curve
 When, plotted as a species accumulation curve over time, it looks linear, like this:
```{r}
 #just making a vector of years from 1920 to 2020
 alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
 colnames(alltheyears) <- "Year"

 #grouping my simulated species by the year they naturalized
 year_count <- as.data.frame(table(sim_spp$Naturalization_Year)) %>%
 # and renaming columns
   dplyr::rename(Year = Var1, Taxa_Naturalized = Freq)

 #merging with all the years
 year_count <-merge(alltheyears, year_count, all = TRUE)
 #making the NA's become 0s, because 0 new records were reported in those years
 year_count[is.na(year_count)] = 0

 #calculating species accumulation over time
 accumulated <- accumulate(year_count$Taxa_Naturalized, sum)
 accumulated <- cbind(year_count, accumulated)
 accumulated
```

```{r}
mean(accumulated$Taxa_Naturalized)
```
```{r}
hist(accumulated$Taxa_Naturalized)
```
```{r}
hist(accumulated$Year)
```

```{r}
 write.csv(sim_spp, file = "sim_spp.csv", row.names = FALSE)
 write.csv(accumulated, file = "accumulated.csv", row.names = FALSE)
```