---
title: "simstudy_PART4_recordretrieval"
author: "Kelsey"
date: "10/2/2021"
output: html_document
---


```{r, }
require(knitr)
```

```{r}
#get the needed packages
if(!require("pacman")){
	install.packages("pacman")
	library(pacman)}
p_load("dplyr", "ggplot2", "viridis", "tidyr",  "fGarch", "sn", "purrr", "forcats", "stats", "changepoint", "changepoint.np")
```

```{r}
sim_spp <- read.csv(file = "sim_spp.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F)
accumulated <- read.csv(file = "accumulated.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F)

accumulated
```

### Changepoint Analysis

```{r}
a2 <- changepoint::cpt.meanvar(accumulated$Taxa_Naturalized, penalty = "MBIC", method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)
```

```{r}
changepoint::plot(a2)
```
```{r}
changepoint::param.est(a2)
```
```{r}
accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
```


# COMPLETENESS PART
# Scenario 10: 60% completeness


## Simulate Hypothetical Lag Times

For this scenario, let's say that all lags have a mean of 10 years

```{r}
sim_spp10 <- sim_spp
```

##### 3.0
Simulate the lag between introduction and naturalization
##### 3.0
Simulate the lag between introduction and naturalization



```{r}
shape = 1.3
rate = 0.15
```

```{r}
set.seed(NULL)
Intro_Nat_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
Intro_Nat_Lag <- subset(Intro_Nat_Lag, Intro_Nat_Lag >=0 & Intro_Nat_Lag <=500)
Intro_Nat_Lag <- as.data.frame(sample(Intro_Nat_Lag$Intro_Nat_Lag, 1000))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
ave(Intro_Nat_Lag)
#plotting histogram
ggplot(Intro_Nat_Lag, aes(x = Intro_Nat_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Intro_Nat_Lag)
```

##### 3.1
Then, simulate the lag between naturalization and detection

```{r}
#shape = 1.3
#rate = 0.15

set.seed(NULL)
Nat_Detect_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate))) *2

colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
Nat_Detect_Lag <- subset(Nat_Detect_Lag, Nat_Detect_Lag >=0 & Nat_Detect_Lag <=500)
Nat_Detect_Lag <- as.data.frame(sample(Nat_Detect_Lag$Nat_Detect_Lag, 1000))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
ave(Nat_Detect_Lag)
#plotting histogram
ggplot(Nat_Detect_Lag, aes(x = Nat_Detect_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Nat_Detect_Lag)

```



third, simulate the lag between vouchering and identification

```{r}

#shape = 1.3
#rate = 0.15

set.seed(NULL)
Vouch_ID_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
Vouch_ID_Lag <- subset(Vouch_ID_Lag, Vouch_ID_Lag >=0 & Vouch_ID_Lag <=500)
Vouch_ID_Lag <- as.data.frame(sample(Vouch_ID_Lag$Vouch_ID_Lag, 1000))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
ave(Vouch_ID_Lag)
#plotting histogram
ggplot(Vouch_ID_Lag, aes(x = Vouch_ID_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Vouch_ID_Lag)
```

##### 3.4

Then, simulate the lag between IDing and reporting
```{r}
#shape = 1.3
#rate = 0.15

set.seed(NULL)
ID_Report_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(ID_Report_Lag) <- "ID_Report_Lag"
ID_Report_Lag <- subset(ID_Report_Lag, ID_Report_Lag >=0 & ID_Report_Lag <=500)
ID_Report_Lag <- as.data.frame(sample(ID_Report_Lag$ID_Report_Lag, 1000))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
ave(ID_Report_Lag)
#plotting histogram
ggplot(ID_Report_Lag, aes(x = ID_Report_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(ID_Report_Lag)
# summary(ID_Report_Lag)
```


##### 3.5
Then, simulate the lag between reporting  and when someone makes a resource that is compiled and formatted
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
#shape = 1.3
#rate = 0.15

set.seed(NULL)
Report_Compile_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
Report_Compile_Lag <- subset(Report_Compile_Lag, Report_Compile_Lag >=0 & Report_Compile_Lag <=500)
Report_Compile_Lag <- as.data.frame(sample(Report_Compile_Lag$Report_Compile_Lag, 1000))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
ave(Report_Compile_Lag)
#plotting histogram
ggplot(Report_Compile_Lag, aes(x = Report_Compile_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Report_Compile_Lag)
# summary(Report_Compile_Lag)
```


##### 3.6 Okay, lets add up the lags
So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag,

All created by adding or subtracting from the naturalization year
```{r}
sim_spp10$Intro_Year <- sim_spp10$Naturalization_Year - Intro_Nat_Lag$Intro_Nat_Lag
sim_spp10$Detect_Year <- sim_spp10$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag
sim_spp10$Voucher_Year <- sim_spp10$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag 
sim_spp10$ID_Year <- sim_spp10$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag  + Vouch_ID_Lag$Vouch_ID_Lag
sim_spp10$Report_Year <- sim_spp10$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag  + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag


sim_spp10$Compile_Year <- sim_spp10$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag  + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag + Report_Compile_Lag$Report_Compile_Lag


sim_spp10$All_Processing_Lags <- Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag
#reordering according to Report Year
sim_spp10 <- sim_spp10 %>% dplyr::select(Species_IDNo, Intro_Year, Naturalization_Year, Detect_Year, Voucher_Year, ID_Year, Report_Year, Compile_Year, All_Processing_Lags)
sim_spp10
```

### Cropping the Dataset because we can't know about species that haven't been reported yet


How many species did we lose that hadn't been reported yet?
```{r}
sim_spp10_crop <- subset(sim_spp10, Report_Year < 2020)
```

```{r}
nrow(sim_spp10) -nrow(sim_spp10_crop)
```
```{r}
retrieval_rate = 0.60
sampled <- as.data.frame(sample(sim_spp10_crop$Species_IDNo, (nrow(sim_spp10_crop)* retrieval_rate)))
colnames(sampled) <- "Species_IDNo"
sim_spp10_crop_sample <- merge(sampled, sim_spp10_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE)
sim_spp10_crop_sample
```


## Step 4. Analyse Trends - A) Voucher Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(sim_spp10_crop_sample$Voucher_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Vouchered = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum)
accumulated_A <- cbind(year_count, accumulated_A)
```

## Step 5. Analyse Trends - B) Report Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp10_crop_sample$Report_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Reported = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_B <- accumulate(year_count$Taxa_Reported, sum)
accumulated_B <- cbind(year_count, accumulated_B)
```
## Step 6. Analyse Trends - C) Intro Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp10_crop_sample$Intro_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Introduced = Freq)
year_count

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_C <- accumulate(year_count$Taxa_Introduced, sum)
accumulated_C <- cbind(year_count, accumulated_C)
```

## Step 6. Analyse Trends - D) Compiled Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp10_crop_sample$Compile_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Compiled = Freq)


#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_D <- accumulate(year_count$Taxa_Compiled, sum)
accumulated_D <- cbind(year_count, accumulated_D)
```

#### Changepoint - Evidence Collection

```{r}
plot(accumulated_A$Taxa_Vouchered)
```
```{r}
a1 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

#plot(a1, diagnostic = TRUE)
```
```{r}
selectedcp = 4
#changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)")
changepoint::plot(a1, ncpts = selectedcp)
```
```{r}
changepoint::cpts.full(a1) + (accumulated_A$Year[1] -1)
```
```{r}
firstrow = 3
secondrow = firstrow +1
a2 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "Manual",  pen.value=((pen.value.full(a1)[firstrow] + pen.value.full(a1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A))
chpoints
chpoints + accumulated_A$Year[1]
segmeans <- changepoint::param.est(a2)$mean
segmeans
segvars <- (changepoint::param.est(a2)$variance)
sqrt(segvars)
```
```{r}
seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1])
seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2])
seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3])
seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4])
seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5])
seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6])
seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7])
seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8])
seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1])
seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2])
seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3])
seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4])
seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5])
seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6])
seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7])
seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8])

segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean)
segwmeans <- segwmeans[!is.na(segwmeans)]
segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar)
segwvars <- segwvars[!is.na(segwvars)]

startptindex = 3
endptindex = 4

print("How many species did we lose that hadn't been reported yet?")
spplostharvest <- nrow(sim_spp10) -nrow(sim_spp10_crop)
spplostharvest

spplostretrieval <- nrow(sim_spp10_crop) - nrow(sim_spp10_crop_sample)
spplostretrieval

print("start seg length, mean, std: ")
chpoints[startptindex]
round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2)
round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2)

print("truncated seg length, mean, std:")
chpoints[endptindex] - chpoints[startptindex]
round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2)
round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2)

print("ending seg length, mean, std:")
nrow(accumulated_A) - chpoints[endptindex]
round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2)
round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2)

print("mean of whole, changepoints")
mean(accumulated_A$Taxa_Vouchered)
chpoints + accumulated_A$Year[1]
```
```{r}
changepoint::plot(a2, type='p')
```



#### Changepoints - Intro Date

```{r}
plot(accumulated_C$Taxa_Introduced)
```
```{r}
c1 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

#plot(c1, diagnostic = TRUE)
```
```{r}
selectedcp = 4
#changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)")
changepoint::plot(c1, ncpts = selectedcp)
```

```{r}
changepoint::cpts.full(c1) + as.numeric(accumulated_C$Year[1]) -1
```
```{r}
firstrow = 5
secondrow = firstrow +1
c2 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "Manual",  pen.value=((pen.value.full(c1)[firstrow] + pen.value.full(c1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

chpoints <- c( 0, changepoint::cpts(c2), nrow(accumulated_C))
chpoints
chpoints + as.numeric(accumulated_C$Year[1])
segmeans <- changepoint::param.est(c2)$mean
segmeans
segvars <- (changepoint::param.est(c2)$variance)
sqrt(segvars)
```
```{r}
seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1])
seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2])
seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3])
seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4])
seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5])
seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6])
seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7])
seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8])
seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1])
seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2])
seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3])
seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4])
seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5])
seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6])
seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7])
seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8])

segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean)
segwmeans <- segwmeans[!is.na(segwmeans)]
segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar)
segwvars <- segwvars[!is.na(segwvars)]

startptindex = 2
endptindex = 4

print("How many species did we lose that hadn't been reported yet?")
spplostharvest <- nrow(sim_spp10) -nrow(sim_spp10_crop)
spplostharvest

spplostretrieval <- nrow(sim_spp10_crop) - nrow(sim_spp10_crop_sample)
spplostretrieval

print("start seg length, mean, std: ")
chpoints[startptindex]
round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2)
round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2)

print("truncated seg length, mean, std:")
chpoints[endptindex] - chpoints[startptindex]
round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2)
round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2)

print("ending seg length, mean, std:")
nrow(accumulated_C) - chpoints[endptindex]
round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2)
round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2)

print("mean of whole, changepoints")
mean(accumulated_C$Taxa_Introduced)
chpoints + as.numeric(accumulated_C$Year[1])
```
```{r}
changepoint::plot(c2, type='p')
```




```{r}
#making into a single dataframe
#accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered)
accumulated_B  <- accumulated_B %>% dplyr::rename(accumulated_species = accumulated_B, No_Taxa = Taxa_Reported)
accumulated_C  <- accumulated_C %>% dplyr::rename(accumulated_species = accumulated_C, No_Taxa = Taxa_Introduced)
accumulated_D  <- accumulated_D %>% dplyr::rename(accumulated_species = accumulated_D, No_Taxa = Taxa_Compiled)
accumulated$Analysis_Type <- "True naturalization date"
accumulated$Order <- 2
accumulated_A$Analysis_Type <- "Detection date"
accumulated_A$Order <- 1
accumulated_B$Analysis_Type <- "Report Date"
accumulated_B$Order <- NA
accumulated_C$Analysis_Type <- "Estimated Introduction Date"
accumulated_C$Order <- 3
accumulated_D$Analysis_Type <- "Compiled Date"
accumulated_D$Order <- NA

accumulated_A_B_C_D <- rbind(accumulated, accumulated_A, accumulated_B, accumulated_C, accumulated_D)
accumulated_A_B_C_D
```

```{r}
# DATA <- accumulated_A_B_C_D %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
# DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
# #plotting total species accumulation over time
# ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(No_Taxa), group = Analysis_Type, color = Analysis_Type)) +
#   geom_smooth( size=1.5, alpha=0.8, level = 0, method = "loess", fullrange=TRUE) +
#   scale_color_viridis(discrete = TRUE) +
#   scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
#   guides(colour = guide_legend(reverse=T)) +
#  theme_classic() +
#   labs(title = "Scenario10 Nat Rate", x = "Year", y = "Number of Species")
# ggsave(filename = "Scenario10_ rate.png", device = "png", scale = 1, width =6, dpi = 600)
```

```{r}
DATA <- accumulated_A_B_C_D # %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, Order))
DATA <- subset(DATA, DATA$Analysis_Type != "Report Date")
DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
poor_ret <- DATA
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) +
  geom_line( size=2, alpha=0.70) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1900, 1920, 1940, 1960, 1980, 2000, 2020)) +
 # guides(colour = guide_legend(reverse=T)) +
 theme_light() +
  theme(legend.position = "none",panel.border = element_blank(),legend.title=element_blank()) +
  labs(title = "E:  60% of Records Retrieved", x = "Year", y = "Total Naturalized Species")
ggsave(filename = "Fig4_60natrecordsretreived.png", device = "png", scale = 1, width =4, dpi = 600)
```


# Scenario 11: 66.67% completeness

## Simulate Hypothetical Lag Times

For this scenario, let's say that all lags have a mean of 10 years

```{r}
sim_spp11 <- sim_spp
```

##### 3.0


##### 3.0
Simulate the lag between introduction and naturalization



```{r}
shape = 1.3
rate = 0.15
```

```{r}
set.seed(NULL)
Intro_Nat_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
Intro_Nat_Lag <- subset(Intro_Nat_Lag, Intro_Nat_Lag >=0 & Intro_Nat_Lag <=500)
Intro_Nat_Lag <- as.data.frame(sample(Intro_Nat_Lag$Intro_Nat_Lag, 1000))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
ave(Intro_Nat_Lag)
#plotting histogram
ggplot(Intro_Nat_Lag, aes(x = Intro_Nat_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Intro_Nat_Lag)
```

##### 3.1
Then, simulate the lag between naturalization and detection

```{r}
#shape = 1.3
#rate = 0.15

set.seed(NULL)
Nat_Detect_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate))) *2

colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
Nat_Detect_Lag <- subset(Nat_Detect_Lag, Nat_Detect_Lag >=0 & Nat_Detect_Lag <=500)
Nat_Detect_Lag <- as.data.frame(sample(Nat_Detect_Lag$Nat_Detect_Lag, 1000))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
ave(Nat_Detect_Lag)
#plotting histogram
ggplot(Nat_Detect_Lag, aes(x = Nat_Detect_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Nat_Detect_Lag)

```

##### 3.3

third, simulate the lag between vouchering and identification

```{r}

#shape = 1.3
#rate = 0.15

set.seed(NULL)
Vouch_ID_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
Vouch_ID_Lag <- subset(Vouch_ID_Lag, Vouch_ID_Lag >=0 & Vouch_ID_Lag <=500)
Vouch_ID_Lag <- as.data.frame(sample(Vouch_ID_Lag$Vouch_ID_Lag, 1000))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
ave(Vouch_ID_Lag)
#plotting histogram
ggplot(Vouch_ID_Lag, aes(x = Vouch_ID_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Vouch_ID_Lag)
```

##### 3.4

Then, simulate the lag between IDing and reporting
```{r}
#shape = 1.3
#rate = 0.15

set.seed(NULL)
ID_Report_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(ID_Report_Lag) <- "ID_Report_Lag"
ID_Report_Lag <- subset(ID_Report_Lag, ID_Report_Lag >=0 & ID_Report_Lag <=500)
ID_Report_Lag <- as.data.frame(sample(ID_Report_Lag$ID_Report_Lag, 1000))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
ave(ID_Report_Lag)
#plotting histogram
ggplot(ID_Report_Lag, aes(x = ID_Report_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(ID_Report_Lag)
# summary(ID_Report_Lag)
```


##### 3.5
Then, simulate the lag between reporting  and when someone makes a resource that is compiled and formatted
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
#shape = 1.3
#rate = 0.15

set.seed(NULL)
Report_Compile_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
Report_Compile_Lag <- subset(Report_Compile_Lag, Report_Compile_Lag >=0 & Report_Compile_Lag <=500)
Report_Compile_Lag <- as.data.frame(sample(Report_Compile_Lag$Report_Compile_Lag, 1000))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
ave(Report_Compile_Lag)
#plotting histogram
ggplot(Report_Compile_Lag, aes(x = Report_Compile_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Report_Compile_Lag)
# summary(Report_Compile_Lag)
```


##### 3.6 Okay, lets add up the lags
So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag,

All created by adding or subtracting from the naturalization year
```{r}
sim_spp11$Intro_Year <- sim_spp11$Naturalization_Year - Intro_Nat_Lag$Intro_Nat_Lag
sim_spp11$Detect_Year <- sim_spp11$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag
sim_spp11$Voucher_Year <- sim_spp11$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag 
sim_spp11$ID_Year <- sim_spp11$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag  + Vouch_ID_Lag$Vouch_ID_Lag
sim_spp11$Report_Year <- sim_spp11$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag  + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag


sim_spp11$Compile_Year <- sim_spp11$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag  + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag + Report_Compile_Lag$Report_Compile_Lag


sim_spp11$All_Processing_Lags <- Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag
#reordering according to Report Year
sim_spp11 <- sim_spp11 %>% dplyr::select(Species_IDNo, Intro_Year, Naturalization_Year, Detect_Year, Voucher_Year, ID_Year, Report_Year, Compile_Year, All_Processing_Lags)
```

### Cropping the Dataset because we can't know about species that haven't been reported yet


How many species did we lose that hadn't been reported yet?
```{r}
sim_spp11_crop <- subset(sim_spp11, Report_Year < 2020)
sim_spp11_crop
```

```{r}
nrow(sim_spp11) -nrow(sim_spp11_crop)
```
```{r}
retrieval_rate = .90
sampled <- as.data.frame(sample(sim_spp11_crop$Species_IDNo, (nrow(sim_spp11_crop)* retrieval_rate)))
colnames(sampled) <- "Species_IDNo"
sim_spp11_crop_sample <- merge(sampled, sim_spp11_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE)
sim_spp11_crop_sample
```

## Step 4. Analyse Trends - A) Voucher Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(sim_spp11_crop_sample$Voucher_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Vouchered = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum)
accumulated_A <- cbind(year_count, accumulated_A)
```

## Step 5. Analyse Trends - B) Report Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp11_crop_sample$Report_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Reported = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_B <- accumulate(year_count$Taxa_Reported, sum)
accumulated_B <- cbind(year_count, accumulated_B)
```
## Step 6. Analyse Trends - C) Intro Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp11_crop_sample$Intro_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Introduced = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_C <- accumulate(year_count$Taxa_Introduced, sum)
accumulated_C <- cbind(year_count, accumulated_C)
```

## Step 6. Analyse Trends - D) Compiled Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp11_crop_sample$Compile_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Compiled = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_D <- accumulate(year_count$Taxa_Compiled, sum)
accumulated_D <- cbind(year_count, accumulated_D)
```
#### Changepoint - Evidence Collection


```{r}
plot(accumulated_A$Taxa_Vouchered)
```
```{r}
a1 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

#plot(a1, diagnostic = TRUE)
```
```{r}
selectedcp = 5
#changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)")
changepoint::plot(a1, ncpts = selectedcp)
```
```{r}
changepoint::cpts.full(a1) + (accumulated_A$Year[1] -1)
```
```{r}
firstrow = 4
secondrow = firstrow +1
a2 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "Manual",  pen.value=((pen.value.full(a1)[firstrow] + pen.value.full(a1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A))
chpoints
chpoints + accumulated_A$Year[1]
segmeans <- changepoint::param.est(a2)$mean
segmeans
segvars <- (changepoint::param.est(a2)$variance)
sqrt(segvars)
```
```{r}
seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1])
seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2])
seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3])
seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4])
seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5])
seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6])
seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7])
seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8])
seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1])
seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2])
seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3])
seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4])
seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5])
seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6])
seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7])
seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8])

segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean)
segwmeans <- segwmeans[!is.na(segwmeans)]
segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar)
segwvars <- segwvars[!is.na(segwvars)]
```
```{r}
startptindex = 3
endptindex = 4

print("How many species did we lose that hadn't been reported yet?")
spplostharvest <- nrow(sim_spp11) -nrow(sim_spp11_crop)
spplostharvest

spplostretrieval <- nrow(sim_spp11_crop) - nrow(sim_spp11_crop_sample)
spplostretrieval

print("start seg length, mean, std: ")
chpoints[startptindex]
round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2)
round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2)

print("truncated seg length, mean, std:")
chpoints[endptindex] - chpoints[startptindex]
round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2)
round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2)

print("ending seg length, mean, std:")
nrow(accumulated_A) - chpoints[endptindex]
round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2)
round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2)
```
```{r}
changepoint::plot(a2, type='p')
```



#### Changepoints - Intro Date

```{r}
plot(accumulated_C$Taxa_Introduced)
```
```{r}
c1 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

#plot(c1, diagnostic = TRUE)
```
```{r}
selectedcp = 5
#changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)")
changepoint::plot(c1, ncpts = selectedcp)
```

```{r}
changepoint::cpts.full(c1) + as.numeric(accumulated_C$Year[1]) -1
```
```{r}
firstrow = 5
secondrow = firstrow +1
c2 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "Manual",  pen.value=((pen.value.full(c1)[firstrow] + pen.value.full(c1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

chpoints <- c( 0, changepoint::cpts(c2), nrow(accumulated_C))
chpoints
chpoints + as.numeric(accumulated_C$Year[1])
segmeans <- changepoint::param.est(c2)$mean
segmeans
segvars <- (changepoint::param.est(c2)$variance)
sqrt(segvars)
```
```{r}
seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1])
seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2])
seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3])
seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4])
seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5])
seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6])
seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7])
seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8])
seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1])
seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2])
seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3])
seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4])
seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5])
seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6])
seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7])
seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8])

segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean)
segwmeans <- segwmeans[!is.na(segwmeans)]
segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar)
segwvars <- segwvars[!is.na(segwvars)]
```
```{r}
startptindex = 1
endptindex = 2

print("How many species did we lose that hadn't been reported yet?")
spplostharvest <- nrow(sim_spp11) -nrow(sim_spp11_crop)
spplostharvest

spplostretrieval <- nrow(sim_spp11_crop) - nrow(sim_spp11_crop_sample)
spplostretrieval

print("start seg length, mean, std: ")
chpoints[startptindex]
round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2)
round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2)

print("truncated seg length, mean, std:")
chpoints[endptindex] - chpoints[startptindex]
round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2)
round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2)

print("ending seg length, mean, std:")
nrow(accumulated_A) - chpoints[endptindex]
round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2)
round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2)
```
```{r}
changepoint::plot(c2, type='p')
```



```{r}
#making into a single dataframe
#accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered)
accumulated_B  <- accumulated_B %>% dplyr::rename(accumulated_species = accumulated_B, No_Taxa = Taxa_Reported)
accumulated_C  <- accumulated_C %>% dplyr::rename(accumulated_species = accumulated_C, No_Taxa = Taxa_Introduced)
accumulated_D  <- accumulated_D %>% dplyr::rename(accumulated_species = accumulated_D, No_Taxa = Taxa_Compiled)
accumulated$Analysis_Type <- "True naturalization date"
accumulated$Order <- 2
accumulated_A$Analysis_Type <- "Detection date"
accumulated_A$Order <- 1
accumulated_B$Analysis_Type <- "Report Date"
accumulated_B$Order <- NA
accumulated_C$Analysis_Type <- "Estimated Introduction Date"
accumulated_C$Order <- 3
accumulated_D$Analysis_Type <- "Compiled Date"
accumulated_D$Order <- NA


accumulated_A_B_C_D <- rbind(accumulated, accumulated_A, accumulated_B, accumulated_C, accumulated_D)
accumulated_A_B_C_D
```

```{r}
# DATA <- accumulated_A_B_C_D %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
# DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
# #plotting total species accumulation over time
# ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(No_Taxa), group = Analysis_Type, color = Analysis_Type)) +
#   geom_smooth( size=1.5, alpha=0.8, level = 0, method = "loess", fullrange=TRUE) +
#   scale_color_viridis(discrete = TRUE) +
#   scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
#   guides(colour = guide_legend(reverse=T)) +
#  theme_classic() +
#   labs(title = "Scenario11 Nat Rate", x = "Year", y = "Number of Species")
# ggsave(filename = "Scenario11_ rate.png", device = "png", scale = 1, width =6, dpi = 600)
```

```{r}
DATA <- accumulated_A_B_C_D # %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, Order))
DATA <- subset(DATA, DATA$Analysis_Type != "Report Date")
DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
decent_ret <- DATA
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) +
  geom_line( size=2, alpha=0.70) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1900, 1920, 1940, 1960, 1980, 2000, 2020)) +
 # guides(colour = guide_legend(reverse=T)) +
 theme_light() +
  theme(legend.position = "none",panel.border = element_blank(),legend.title=element_blank()) +
  labs(title = "D:  90% of Records Retrieved", x = "Year", y = "Total Naturalized Species")
ggsave(filename = "Fig4_90natrecordretreived.png", device = "png", scale = 1, width =4, dpi = 600)
```





# REAL-WOLRD MIXED EFFECTS PART


# Scenario 12: Mixed data: 60% retrieval from 1920-1970, 90% from 1970 - 2020

## Simulate Hypothetical Lag Times

For this scenario, let's say that all lags have a mean of 10 years

```{r}
sim_spp12 <- sim_spp
sim_spp12
```


##### 3.0
Simulate the lag between introduction and naturalization



```{r}
shape = 1.3
rate = 0.15
```

```{r}
set.seed(NULL)
Intro_Nat_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
Intro_Nat_Lag <- subset(Intro_Nat_Lag, Intro_Nat_Lag >=0 & Intro_Nat_Lag <=500)
Intro_Nat_Lag <- as.data.frame(sample(Intro_Nat_Lag$Intro_Nat_Lag, 1000))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
ave(Intro_Nat_Lag)
#plotting histogram
ggplot(Intro_Nat_Lag, aes(x = Intro_Nat_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Intro_Nat_Lag)
```

##### 3.1
Then, simulate the lag between naturalization and detection

```{r}
#shape = 1.3
#rate = 0.15

set.seed(NULL)
Nat_Detect_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate))) *2

colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
Nat_Detect_Lag <- subset(Nat_Detect_Lag, Nat_Detect_Lag >=0 & Nat_Detect_Lag <=500)
Nat_Detect_Lag <- as.data.frame(sample(Nat_Detect_Lag$Nat_Detect_Lag, 1000))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
ave(Nat_Detect_Lag)
#plotting histogram
ggplot(Nat_Detect_Lag, aes(x = Nat_Detect_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Nat_Detect_Lag)

```


##### 3.3

third, simulate the lag between vouchering and identification

```{r}

#shape = 1.3
#rate = 0.15

set.seed(NULL)
Vouch_ID_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
Vouch_ID_Lag <- subset(Vouch_ID_Lag, Vouch_ID_Lag >=0 & Vouch_ID_Lag <=500)
Vouch_ID_Lag <- as.data.frame(sample(Vouch_ID_Lag$Vouch_ID_Lag, 1000))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
ave(Vouch_ID_Lag)
#plotting histogram
ggplot(Vouch_ID_Lag, aes(x = Vouch_ID_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Vouch_ID_Lag)
```

##### 3.4

Then, simulate the lag between IDing and reporting
```{r}
#shape = 1.3
#rate = 0.15

set.seed(NULL)
ID_Report_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(ID_Report_Lag) <- "ID_Report_Lag"
ID_Report_Lag <- subset(ID_Report_Lag, ID_Report_Lag >=0 & ID_Report_Lag <=500)
ID_Report_Lag <- as.data.frame(sample(ID_Report_Lag$ID_Report_Lag, 1000))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
ave(ID_Report_Lag)
#plotting histogram
ggplot(ID_Report_Lag, aes(x = ID_Report_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(ID_Report_Lag)
# summary(ID_Report_Lag)
```


##### 3.5
Then, simulate the lag between reporting  and when someone makes a resource that is compiled and formatted
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
#shape = 1.3
#rate = 0.15

set.seed(NULL)
Report_Compile_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
Report_Compile_Lag <- subset(Report_Compile_Lag, Report_Compile_Lag >=0 & Report_Compile_Lag <=500)
Report_Compile_Lag <- as.data.frame(sample(Report_Compile_Lag$Report_Compile_Lag, 1000))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
ave(Report_Compile_Lag)
#plotting histogram
ggplot(Report_Compile_Lag, aes(x = Report_Compile_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Report_Compile_Lag)
# summary(Report_Compile_Lag)
```


##### 3.6 Okay, lets add up the lags
So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag,

All created by adding or subtracting from the naturalization year
```{r}
sim_spp12$Intro_Year <- sim_spp12$Naturalization_Year - Intro_Nat_Lag$Intro_Nat_Lag
sim_spp12$Detect_Year <- sim_spp12$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag
sim_spp12$Voucher_Year <- sim_spp12$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag 
sim_spp12$ID_Year <- sim_spp12$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag  + Vouch_ID_Lag$Vouch_ID_Lag
sim_spp12$Report_Year <- sim_spp12$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag  + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag


sim_spp12$Compile_Year <- sim_spp12$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag  + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag + Report_Compile_Lag$Report_Compile_Lag


sim_spp12$All_Processing_Lags <- Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag
#reordering according to Report Year
sim_spp12 <- sim_spp12 %>% dplyr::select(Species_IDNo, Intro_Year, Naturalization_Year, Detect_Year, Voucher_Year, ID_Year, Report_Year, Compile_Year, All_Processing_Lags)
```

### Cropping the Dataset because we can't know about species that haven't been reported yet

How many species did we lose that hadn't been reported yet?
```{r}
sim_spp12_crop <- subset(sim_spp12, Report_Year < 2020)
sim_spp12_crop
```

```{r}
nrow(sim_spp12) -nrow(sim_spp12_crop)
```
```{r}
sim_spp12_crop_early <- subset(sim_spp12_crop, Voucher_Year < 1970)
retrieval_rate = 0.60
sampled_early <- as.data.frame(sample(sim_spp12_crop_early$Species_IDNo, (nrow(sim_spp12_crop_early)* retrieval_rate)))
colnames(sampled_early) <- "Species_IDNo"
sampled_early
```


```{r}
sim_spp12_crop_late <- subset(sim_spp12_crop, Voucher_Year > 1970)
retrieval_rate = 0.90
sampled_late <- as.data.frame(sample(sim_spp12_crop_late$Species_IDNo, (nrow(sim_spp12_crop_late)* retrieval_rate)))
colnames(sampled_late) <- "Species_IDNo"
sampled_late
```


```{r}
sampled <- rbind(sampled_early, sampled_late)
sim_spp12_crop_sample <- merge(sampled, sim_spp12_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE)
sim_spp12_crop_sample
```


## Step 4. Analyse Trends - A) Voucher Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(sim_spp12_crop_sample$Voucher_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Vouchered = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum)
accumulated_A <- cbind(year_count, accumulated_A)
```

## Step 5. Analyse Trends - B) Report Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp12_crop_sample$Report_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Reported = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_B <- accumulate(year_count$Taxa_Reported, sum)
accumulated_B <- cbind(year_count, accumulated_B)
```
## Step 6. Analyse Trends - C) Intro Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp12_crop_sample$Intro_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Introduced = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_C <- accumulate(year_count$Taxa_Introduced, sum)
accumulated_C <- cbind(year_count, accumulated_C)
```

## Step 6. Analyse Trends - D) Compiled Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp12_crop_sample$Compile_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Compiled = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_D <- accumulate(year_count$Taxa_Compiled, sum)
accumulated_D <- cbind(year_count, accumulated_D)
```
#### Changepoint - Evidence Collection


```{r}
plot(accumulated_A$Taxa_Vouchered)
```
```{r}
a1 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

#plot(a1, diagnostic = TRUE)
```
```{r}
selectedcp = 4
#changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)")
changepoint::plot(a1, ncpts = selectedcp)
```
```{r}
changepoint::cpts.full(a1) + (accumulated_A$Year[1] -1)
```
```{r}
firstrow = 4
secondrow = firstrow +1
a2 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "Manual",  pen.value=((pen.value.full(a1)[firstrow] + pen.value.full(a1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A))
chpoints
chpoints + accumulated_A$Year[1]
segmeans <- changepoint::param.est(a2)$mean
segmeans
segvars <- (changepoint::param.est(a2)$variance)
sqrt(segvars)
```
```{r}
seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1])
seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2])
seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3])
seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4])
seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5])
seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6])
seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7])
seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8])
seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1])
seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2])
seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3])
seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4])
seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5])
seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6])
seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7])
seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8])

segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean)
segwmeans <- segwmeans[!is.na(segwmeans)]
segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar)
segwvars <- segwvars[!is.na(segwvars)]
```
```{r}
startptindex = 3
endptindex = 4

print("How many species did we lose that hadn't been reported yet?")
spplostharvest <- nrow(sim_spp12) -nrow(sim_spp12_crop)
spplostharvest

spplostretrieval <- nrow(sim_spp12_crop) - nrow(sim_spp12_crop_sample)
spplostretrieval

print("start seg length, mean, std: ")
chpoints[startptindex]
round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2)
round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2)

print("truncated seg length, mean, std:")
chpoints[endptindex] - chpoints[startptindex]
round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2)
round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2)

print("ending seg length, mean, std:")
nrow(accumulated_A) - chpoints[endptindex]
round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2)
round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2)
```
```{r}
changepoint::plot(a2, type='p')
```



#### Changepoints - Intro Date

```{r}
plot(accumulated_C$Taxa_Introduced)
```
```{r}
c1 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

#plot(c1, diagnostic = TRUE)
```
```{r}
selectedcp = 4
#changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)")
changepoint::plot(c1, ncpts = selectedcp)
```

```{r}
changepoint::cpts.full(c1) + as.numeric(accumulated_C$Year[1]) -1
```
```{r}
firstrow = 5
secondrow = firstrow +1
c2 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "Manual",  pen.value=((pen.value.full(c1)[firstrow] + pen.value.full(c1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

chpoints <- c( 0, changepoint::cpts(c2), nrow(accumulated_C))
chpoints
chpoints + as.numeric(accumulated_C$Year[1])
segmeans <- changepoint::param.est(c2)$mean
segmeans
segvars <- (changepoint::param.est(c2)$variance)
sqrt(segvars)
```
```{r}
seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1])
seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2])
seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3])
seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4])
seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5])
seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6])
seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7])
seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8])
seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1])
seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2])
seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3])
seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4])
seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5])
seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6])
seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7])
seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8])

segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean)
segwmeans <- segwmeans[!is.na(segwmeans)]
segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar)
segwvars <- segwvars[!is.na(segwvars)]
```
```{r}
startptindex = 1
endptindex = 2

print("How many species did we lose that hadn't been reported yet?")
spplostharvest <- nrow(sim_spp12) -nrow(sim_spp12_crop)
spplostharvest

spplostretrieval <- nrow(sim_spp12_crop) - nrow(sim_spp12_crop_sample)
spplostretrieval

print("start seg length, mean, std: ")
chpoints[startptindex]
round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2)
round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2)

print("truncated seg length, mean, std:")
chpoints[endptindex] - chpoints[startptindex]
round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2)
round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2)

print("ending seg length, mean, std:")
nrow(accumulated_A) - chpoints[endptindex]
round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2)
round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2)
```
```{r}
changepoint::plot(c2, type='p')
```



```{r}
#making into a single dataframe
#accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered)
accumulated_B  <- accumulated_B %>% dplyr::rename(accumulated_species = accumulated_B, No_Taxa = Taxa_Reported)
accumulated_C  <- accumulated_C %>% dplyr::rename(accumulated_species = accumulated_C, No_Taxa = Taxa_Introduced)
accumulated_D  <- accumulated_D %>% dplyr::rename(accumulated_species = accumulated_D, No_Taxa = Taxa_Compiled)
accumulated$Analysis_Type <- "True naturalization date"
accumulated$Order <- 2
accumulated_A$Analysis_Type <- "Detection date"
accumulated_A$Order <- 1
accumulated_B$Analysis_Type <- "Report Date"
accumulated_B$Order <- NA
accumulated_C$Analysis_Type <- "Estimated Introduction Date"
accumulated_C$Order <- 3
accumulated_D$Analysis_Type <- "Compiled Date"
accumulated_D$Order <- NA


accumulated_A_B_C_D <- rbind(accumulated, accumulated_A, accumulated_B, accumulated_C, accumulated_D)
accumulated_A_B_C_D
```

```{r}
# DATA <- accumulated_A_B_C_D %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
# DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
# #plotting total species accumulation over time
# ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(No_Taxa), group = Analysis_Type, color = Analysis_Type)) +
#   geom_smooth( size=1.5, alpha=0.8, level = 0, method = "loess", fullrange=TRUE) +
#   scale_color_viridis(discrete = TRUE) +
#   scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
#   guides(colour = guide_legend(reverse=T)) +
#  theme_classic() +
#   labs(title = "Scenario11 Nat Rate", x = "Year", y = "Number of Species")
# ggsave(filename = "Scenario11_ rate.png", device = "png", scale = 1, width =6, dpi = 600)
```

```{r}
DATA <- accumulated_A_B_C_D #%>% mutate(Analysis_Type = fct_reorder(Analysis_Type, Order))
DATA <- subset(DATA, DATA$Analysis_Type != "Report Date")
DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
lowearly_ret <- DATA
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) +
  geom_line( size=2, alpha=0.70) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1900, 1920, 1940, 1960, 1980, 2000, 2020)) +
 # guides(colour = guide_legend(reverse=T)) +
 theme_light() +
  theme(legend.position = "none",panel.border = element_blank(),legend.title=element_blank()) +
  labs(title = "F:  Higher Record Retrieval \n in Recent Years", x = "Year", y = "Total Naturalized Species")
ggsave(filename = "Fig4_lowearlyhighlater.png", device = "png", scale = 1, width =4, dpi = 600)
```


# Scenario 1??: Mixed data: 90% retrieval from 1920-1970, 60% from 1970 - 2020

## Simulate Hypothetical Lag Times

For this scenario, let's say that all lags have a mean of 10 years

```{r}
sim_spp13 <- sim_spp
sim_spp13
```


##### 3.0
Simulate the lag between introduction and naturalization

```{r}
shape = 1.3
rate = 0.15
```

```{r}
set.seed(NULL)
Intro_Nat_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
Intro_Nat_Lag <- subset(Intro_Nat_Lag, Intro_Nat_Lag >=0 & Intro_Nat_Lag <=500)
Intro_Nat_Lag <- as.data.frame(sample(Intro_Nat_Lag$Intro_Nat_Lag, 1000))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
ave(Intro_Nat_Lag)
#plotting histogram
ggplot(Intro_Nat_Lag, aes(x = Intro_Nat_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Intro_Nat_Lag)
```

##### 3.1
Then, simulate the lag between naturalization and detection

```{r}
#shape = 1.3
#rate = 0.15

set.seed(NULL)
Nat_Detect_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate))) *2

colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
Nat_Detect_Lag <- subset(Nat_Detect_Lag, Nat_Detect_Lag >=0 & Nat_Detect_Lag <=500)
Nat_Detect_Lag <- as.data.frame(sample(Nat_Detect_Lag$Nat_Detect_Lag, 1000))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
ave(Nat_Detect_Lag)
#plotting histogram
ggplot(Nat_Detect_Lag, aes(x = Nat_Detect_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Nat_Detect_Lag)

```

##### 3.3

third, simulate the lag between vouchering and identification

```{r}

#shape = 1.3
#rate = 0.15

set.seed(NULL)
Vouch_ID_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
Vouch_ID_Lag <- subset(Vouch_ID_Lag, Vouch_ID_Lag >=0 & Vouch_ID_Lag <=500)
Vouch_ID_Lag <- as.data.frame(sample(Vouch_ID_Lag$Vouch_ID_Lag, 1000))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
ave(Vouch_ID_Lag)
#plotting histogram
ggplot(Vouch_ID_Lag, aes(x = Vouch_ID_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Vouch_ID_Lag)
```

##### 3.4

Then, simulate the lag between IDing and reporting
```{r}
#shape = 1.3
#rate = 0.15

set.seed(NULL)
ID_Report_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(ID_Report_Lag) <- "ID_Report_Lag"
ID_Report_Lag <- subset(ID_Report_Lag, ID_Report_Lag >=0 & ID_Report_Lag <=500)
ID_Report_Lag <- as.data.frame(sample(ID_Report_Lag$ID_Report_Lag, 1000))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
ave(ID_Report_Lag)
#plotting histogram
ggplot(ID_Report_Lag, aes(x = ID_Report_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(ID_Report_Lag)
# summary(ID_Report_Lag)
```


##### 3.5
Then, simulate the lag between reporting  and when someone makes a resource that is compiled and formatted
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
#shape = 1.3
#rate = 0.15

set.seed(NULL)
Report_Compile_Lag <- as.data.frame(round(rgamma(n = 10000, shape = shape, rate = rate)))

colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
Report_Compile_Lag <- subset(Report_Compile_Lag, Report_Compile_Lag >=0 & Report_Compile_Lag <=500)
Report_Compile_Lag <- as.data.frame(sample(Report_Compile_Lag$Report_Compile_Lag, 1000))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
ave(Report_Compile_Lag)
#plotting histogram
ggplot(Report_Compile_Lag, aes(x = Report_Compile_Lag)) +
  geom_histogram(binwidth = 1) +
theme_light()
summary(Report_Compile_Lag)
# summary(Report_Compile_Lag)
```

##### 3.6 Okay, lets add up the lags
So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag,

All created by adding or subtracting from the naturalization year
```{r}
sim_spp13$Intro_Year <- sim_spp13$Naturalization_Year - Intro_Nat_Lag$Intro_Nat_Lag
sim_spp13$Detect_Year <- sim_spp13$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag
sim_spp13$Voucher_Year <- sim_spp13$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag 
sim_spp13$ID_Year <- sim_spp13$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag  + Vouch_ID_Lag$Vouch_ID_Lag
sim_spp13$Report_Year <- sim_spp13$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag  + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag


sim_spp13$Compile_Year <- sim_spp13$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag  + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag + Report_Compile_Lag$Report_Compile_Lag


sim_spp13$All_Processing_Lags <- Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag
#reordering according to Report Year
sim_spp13 <- sim_spp13 %>% dplyr::select(Species_IDNo, Intro_Year, Naturalization_Year, Detect_Year, Voucher_Year, ID_Year, Report_Year, Compile_Year, All_Processing_Lags)
```

### Cropping the Dataset because we can't know about species that haven't been reported yet


How many species did we lose that hadn't been reported yet?
```{r}
sim_spp13_crop <- subset(sim_spp13, Report_Year < 2020)
sim_spp13_crop
```

```{r}
nrow(sim_spp13) -nrow(sim_spp13_crop)
```
```{r}
sim_spp13_crop_early <- subset(sim_spp13_crop, Voucher_Year < 1970)
retrieval_rate = 0.90
sampled_early <- as.data.frame(sample(sim_spp13_crop_early$Species_IDNo, (nrow(sim_spp13_crop_early)* retrieval_rate)))
colnames(sampled_early) <- "Species_IDNo"
sampled_early
```


```{r}
sim_spp13_crop_late <- subset(sim_spp13_crop, Voucher_Year > 1970)
retrieval_rate = 0.60
sampled_late <- as.data.frame(sample(sim_spp13_crop_late$Species_IDNo, (nrow(sim_spp13_crop_late)* retrieval_rate)))
colnames(sampled_late) <- "Species_IDNo"
sampled_late
```


```{r}
sampled <- rbind(sampled_early, sampled_late)
sim_spp13_crop_sample <- merge(sampled, sim_spp13_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE)
sim_spp13_crop_sample
```


## Step 4. Analyse Trends - A) Voucher Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(sim_spp13_crop_sample$Voucher_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Vouchered = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum)
accumulated_A <- cbind(year_count, accumulated_A)
```

## Step 5. Analyse Trends - B) Report Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp13_crop_sample$Report_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Reported = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_B <- accumulate(year_count$Taxa_Reported, sum)
accumulated_B <- cbind(year_count, accumulated_B)
```
## Step 6. Analyse Trends - C) Intro Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp13_crop_sample$Intro_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Introduced = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_C <- accumulate(year_count$Taxa_Introduced, sum)
accumulated_C <- cbind(year_count, accumulated_C)
```

## Step 6. Analyse Trends - D) Compiled Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp13_crop_sample$Compile_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Compiled = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_D <- accumulate(year_count$Taxa_Compiled, sum)
accumulated_D <- cbind(year_count, accumulated_D)
```
#### Changepoint - Evidence Collection


```{r}
plot(accumulated_A$Taxa_Vouchered)
```
```{r}
a1 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

#plot(a1, diagnostic = TRUE)
```
```{r}
selectedcp = 4
#changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)")
changepoint::plot(a1, ncpts = selectedcp)
```
```{r}
changepoint::cpts.full(a1) + (accumulated_A$Year[1] -1)
```
```{r}
firstrow = 4
secondrow = firstrow +1
a2 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "Manual",  pen.value=((pen.value.full(a1)[firstrow] + pen.value.full(a1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A))
chpoints
chpoints + accumulated_A$Year[1]
segmeans <- changepoint::param.est(a2)$mean
segmeans
segvars <- (changepoint::param.est(a2)$variance)
sqrt(segvars)
```
```{r}
seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1])
seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2])
seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3])
seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4])
seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5])
seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6])
seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7])
seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8])
seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1])
seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2])
seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3])
seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4])
seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5])
seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6])
seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7])
seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8])

segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean)
segwmeans <- segwmeans[!is.na(segwmeans)]
segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar)
segwvars <- segwvars[!is.na(segwvars)]
```
```{r}
startptindex = 3
endptindex = 4

print("How many species did we lose that hadn't been reported yet?")
spplostharvest <- nrow(sim_spp13) -nrow(sim_spp13_crop)
spplostharvest

spplostretrieval <- nrow(sim_spp13_crop) - nrow(sim_spp13_crop_sample)
spplostretrieval

print("start seg length, mean, std: ")
chpoints[startptindex]
round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2)
round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2)

print("truncated seg length, mean, std:")
chpoints[endptindex] - chpoints[startptindex]
round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2)
round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2)

print("ending seg length, mean, std:")
nrow(accumulated_A) - chpoints[endptindex]
round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2)
round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2)
```
```{r}
changepoint::plot(a2, type='p')
```



#### Changepoints - Intro Date

```{r}
plot(accumulated_C$Taxa_Introduced)
```
```{r}
c1 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

#plot(c1, diagnostic = TRUE)
```
```{r}
selectedcp = 4
#changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)")
changepoint::plot(c1, ncpts = selectedcp)
```

```{r}
changepoint::cpts.full(c1) + as.numeric(accumulated_C$Year[1]) -1
```
```{r}
firstrow = 5
secondrow = firstrow +1
c2 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "Manual",  pen.value=((pen.value.full(c1)[firstrow] + pen.value.full(c1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

chpoints <- c( 0, changepoint::cpts(c2), nrow(accumulated_C))
chpoints
chpoints + as.numeric(accumulated_C$Year[1])
segmeans <- changepoint::param.est(c2)$mean
segmeans
segvars <- (changepoint::param.est(c2)$variance)
sqrt(segvars)
```
```{r}
seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1])
seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2])
seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3])
seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4])
seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5])
seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6])
seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7])
seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8])
seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1])
seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2])
seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3])
seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4])
seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5])
seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6])
seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7])
seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8])

segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean)
segwmeans <- segwmeans[!is.na(segwmeans)]
segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar)
segwvars <- segwvars[!is.na(segwvars)]
```
```{r}
startptindex = 1
endptindex = 2

print("How many species did we lose that hadn't been reported yet?")
spplostharvest <- nrow(sim_spp13) -nrow(sim_spp13_crop)
spplostharvest

spplostretrieval <- nrow(sim_spp13_crop) - nrow(sim_spp13_crop_sample)
spplostretrieval

print("start seg length, mean, std: ")
chpoints[startptindex]
round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2)
round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2)

print("truncated seg length, mean, std:")
chpoints[endptindex] - chpoints[startptindex]
round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2)
round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2)

print("ending seg length, mean, std:")
nrow(accumulated_A) - chpoints[endptindex]
round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2)
round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2)
```
```{r}
changepoint::plot(c2, type='p')
```



```{r}
#making into a single dataframe
#accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered)
accumulated_B  <- accumulated_B %>% dplyr::rename(accumulated_species = accumulated_B, No_Taxa = Taxa_Reported)
accumulated_C  <- accumulated_C %>% dplyr::rename(accumulated_species = accumulated_C, No_Taxa = Taxa_Introduced)
accumulated_D  <- accumulated_D %>% dplyr::rename(accumulated_species = accumulated_D, No_Taxa = Taxa_Compiled)
accumulated$Analysis_Type <- "True naturalization date"
accumulated$Order <- 2
accumulated_A$Analysis_Type <- "Detection date"
accumulated_A$Order <- 1
accumulated_B$Analysis_Type <- "Report Date"
accumulated_B$Order <- NA
accumulated_C$Analysis_Type <- "Estimated Introduction Date"
accumulated_C$Order <- 3
accumulated_D$Analysis_Type <- "Compiled Date"
accumulated_D$Order <- NA


accumulated_A_B_C_D <- rbind(accumulated, accumulated_A, accumulated_B, accumulated_C, accumulated_D)
accumulated_A_B_C_D
```

```{r}
# DATA <- accumulated_A_B_C_D %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
# DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
# #plotting total species accumulation over time
# ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(No_Taxa), group = Analysis_Type, color = Analysis_Type)) +
#   geom_smooth( size=1.5, alpha=0.8, level = 0, method = "loess", fullrange=TRUE) +
#   scale_color_viridis(discrete = TRUE) +
#   scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
#   guides(colour = guide_legend(reverse=T)) +
#  theme_classic() +
#   labs(title = "Scenario11 Nat Rate", x = "Year", y = "Number of Species")
# ggsave(filename = "Scenario11_ rate.png", device = "png", scale = 1, width =6, dpi = 600)
```

```{r}
DATA <- accumulated_A_B_C_D # %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, Order))
DATA <- subset(DATA, DATA$Analysis_Type != "Report Date")
DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
highearly_ret <- DATA
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) +
  geom_line( size=2, alpha=0.70) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1900, 1920, 1940, 1960, 1980, 2000, 2020)) +
 # guides(colour = guide_legend(reverse=T)) +
 theme_light() +
  theme(legend.position = "none",panel.border = element_blank(),legend.title=element_blank()) +
  labs(title = "G:  Lower Record Retrieval \n in Recent Years", x = "Year", y = "Total Naturalized Species")
ggsave(filename = "Fig4_highearlylowlater.png", device = "png", scale = 1, width =4, dpi = 600)
```
50% retrieval from 1920-1970, 85% from 1970 - 2020


```{r}
truenat <- subset(decent_ret, decent_ret$Analysis_Type == "True naturalization date")
truenat$Scenario <- "True naturalization date"

decent_ret$Scenario <- "Even retrieval 90% of records"
poor_ret$Scenario <- "Even retrieval 60% of records"
lowearly_ret$Scenario <- "Better retrieval in recent years"
highearly_ret$Scenario <- "Better retrieval in early years"


decent_ret <- subset(decent_ret, decent_ret$Analysis_Type == "Detection date")
poor_ret <- subset(poor_ret, poor_ret$Analysis_Type == "Detection date")
lowearly_ret <- subset(lowearly_ret, lowearly_ret$Analysis_Type == "Detection date")
highearly_ret <- subset(highearly_ret, highearly_ret$Analysis_Type == "Detection date")

DATA <- rbind(truenat, decent_ret, poor_ret, lowearly_ret, highearly_ret)
DATA
```


## FINAL FIGURE
```{r}
library(viridisLite)
library(forcats)
library(ggplot2)

## -- 1  Desired legend / plotting order -------------------------------------
my_order <- c("True naturalization date",
              "Even retrieval 90% of records",
              "Even retrieval 60% of records",
              "Better retrieval in recent years",
              "Better retrieval in early years")

## Make sure the factor follows that order (optional but convenient)
DATA$Scenario <- fct_relevel(DATA$Scenario, my_order)

## -- 2  Helper to drop the brightest colours (unchanged from earlier) -------
drop_yellowish <- function(col_vec, n_keep) {
  n_keep <- max(1, min(n_keep, length(col_vec)))
  rgb_vals   <- col2rgb(col_vec)
  brightness <- colMeans(rgb_vals)
  keep_idx   <- order(brightness)[seq_len(n_keep)]
  col_vec[keep_idx]
}

## -- 3  Build the palette ----------------------------------------------------
n_other   <- length(my_order) - 1                        # colours *besides* black
raw_cols  <- viridis(n_other + 1, option = "D")          # oversample a bit
vir_cols  <- drop_yellowish(raw_cols, n_other)           # keep nicest n_other
named_cols <- c("True naturalization date" = "gray46",
                setNames(vir_cols, my_order[-1]))

## -- 4  Plot -----------------------------------------------------------------
ggplot(DATA,
       aes(x = as.integer(Year), y = accumulated_species,
           colour = Scenario, group = Scenario)) +
  geom_line(size = 1.5, alpha=0.75) +
  scale_colour_manual(breaks = my_order, values = named_cols) +
 scale_x_continuous(breaks = c(1900,1920, 1940, 1960, 1980, 2000, 2020)) +
 theme_light() +
  theme(legend.title=element_blank(), legend.position="right", panel.border = element_blank()) +
  labs(title = "", x = "Year", y = "Total naturalized species")
ggsave(filename = "FINAL_Compare_Retrieval_2025.png", device = "png", scale = 1, width =7, dpi = 600)

```






