---
title: "simstudy_PART6_HS_KBcompare"
author: "Kelsey"
date: "10/5/2021"
output: html_document
---

```{r, }
require(knitr)
```
```{r}
#get the needed packages
if(!require("pacman")){
	install.packages("pacman")
	library(pacman)}
p_load("dplyr", "ggplot2", "viridis", "tidyr",  "fGarch", "sn", "purrr", "forcats", "stats", "changepoint", "changepoint.np", "taxize", "data.table", "tibble", "ggridges", "aomisc")
```

## Get Seebens' Data 

```{r}
HS_data <- read.csv(file = "PacificIslands-Seebens.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F)
HS_data
```

## & Seebens' data with Hawaii replaced with my own data

```{r}
HS_KB_data <- read.csv(file = "PacificIslands-Seebens_HawaiiKB.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F)
HS_KB_data
```
# Sync Taxonomy


<!-- ## Making Sure the Names are in IPNI -->
<!-- ```{r} -->
<!-- gnr.long_HS <- HS_data$NewName %>% -->
<!--   gnr_resolve(data_source_ids = 167, best_match_only = TRUE, with_canonical_ranks = TRUE) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- gnr.long_HS_KB <- HS_KB_data$NewName %>% -->
<!--   gnr_resolve(data_source_ids = 167, best_match_only = TRUE, with_canonical_ranks = TRUE) -->
<!-- ``` -->


<!-- ## Conform to a Single Accepted Name According to Kew's Plants of the World -->

<!-- ```{r} -->
<!-- pow.output_HS <- get_pow(gnr.long_HS$matched_name2, db = "pow", accepted = TRUE, ask = TRUE) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- pow.output_HS <- as.data.frame(pow.output_HS) -->
<!-- inputID_HS <- as.data.frame(cbind(gnr.long_HS$user_supplied_name, gnr.long_HS$matched_name2, pow.output_HS$ids)) -->
<!-- colnames(inputID_HS) <- c("user_supplied_name", "matched_name2", "IPNI_ID") -->
<!-- head(inputID_HS) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- pow.output_HS_KB <- get_pow(gnr.long_HS_KB$matched_name2, db = "pow", accepted = TRUE, ask = TRUE) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- pow.output_HS_KB <- as.data.frame(pow.output_HS_KB) -->
<!-- inputID_HS_KB <- as.data.frame(cbind(gnr.long_HS_KB$user_supplied_name, gnr.long_HS_KB$matched_name2, pow.output_HS_KB$ids)) -->
<!-- colnames(inputID_HS_KB) <- c("user_supplied_name", "matched_name2", "IPNI_ID") -->
<!-- head(inputID_HS_KB) -->
<!-- ``` -->



<!-- ## Separate the Troublemakers (just for now) -->

<!-- ```{r} -->
<!-- inputID_HS.NAsdrop <- subset(inputID_HS, (!is.na(inputID_HS$IPNI_ID))) -->
<!-- nrow(inputID_HS.NAsdrop) -->

<!-- inputID_HS_KB.NAsdrop <- subset(inputID_HS_KB, (!is.na(inputID_HS_KB$IPNI_ID))) -->
<!-- nrow(inputID_HS_KB.NAsdrop) -->

<!-- write.csv(inputID_HS.NAsdrop, "inputID_HS.NAsdrop.csv", row.names = FALSE, quote = FALSE) -->
<!-- write.csv(inputID_HS_KB.NAsdrop, "inputID_HS_KB.NAsdrop.csv", row.names = FALSE, quote = FALSE) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- inputID_HS.NAsonly <- subset(inputID_HS, (is.na(inputID_HS$IPNI_ID))) -->
<!-- nrow(inputID_HS.NAsonly) -->

<!-- inputID_HS_KB.NAsonly <- subset(inputID_HS_KB, (is.na(inputID_HS_KB$IPNI_ID))) -->
<!-- nrow(inputID_HS_KB.NAsonly) -->

<!-- write.csv(inputID_HS.NAsonly, "inputID_HS.NAsonly.csv", row.names = FALSE, quote = FALSE) -->
<!-- write.csv(inputID_HS_KB.NAsonly, "inputID_HS_KB.NAsonly.csv", row.names = FALSE, quote = FALSE) -->
<!-- ``` -->

<!-- ## Get taxon name using IPNI IDs -->

<!-- ```{r} -->
<!-- Get_taxonname <- function(spnames) { -->
<!--   temp <- c() -->
<!--   temp <- pow_lookup(spnames) -->
<!--   ifelse(is.null(temp$meta$name) == "TRUE", "ERROR", temp$meta$name) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- POW.taxonname_HS <- sapply(inputID_HS.NAsdrop$IPNI_ID[], Get_taxonname) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- POW.taxonname_HS_KB <- sapply(inputID_HS_KB.NAsdrop$IPNI_ID[], Get_taxonname) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- POWAcc_HS <- cbind(inputID_HS.NAsdrop, POW.taxonname_HS) -->
<!-- POWAcc_HS_KB <- cbind(inputID_HS_KB.NAsdrop, POW.taxonname_HS_KB) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- write.csv(POWAcc_HS, "POWAcc_HS.csv", row.names = FALSE, quote = FALSE) -->
<!-- write.csv(POWAcc_HS_KB, "POWAcc_HS_KB.csv", row.names = FALSE, quote = FALSE) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- HS_AccNames <- read.csv(file = "POWAcc_HS_fixed.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F) -->
<!-- head(HS_AccNames) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- HS_KB_AccNames <- read.csv(file = "POWAcc_HS_KB_fixed.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F) -->
<!-- head(HS_KB_AccNames) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- HS_data <- merge(HS_AccNames, HS_data, by.x = "user_supplied_name", by.y = "NewName", all.y = TRUE) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- HS_KB_data <- merge(HS_KB_AccNames, HS_KB_data, by.x = "user_supplied_name", by.y = "NewName", all.y = TRUE) -->
<!-- HS_KB_data -->
<!-- ``` -->



<!-- ## Let's compare trends for Hawaii Only -->

<!-- ### Subsetting both datasets to Hawaii Only -->

<!-- ```{r} -->
<!-- HS_data_HI <- subset(HS_data, Region == "Hawaiian Islands") -->
<!-- nrow(HS_data_HI) -->
<!-- HS_KB_data_HI <- subset(HS_KB_data, Region == "Hawaiian Islands") -->
<!-- nrow(HS_KB_data_HI) -->
<!-- ``` -->
<!-- ### Make sure there are no duplicates (if there are, we should take only the first record) -->

<!-- ```{r} -->
<!-- HS_data_HI <- dplyr::arrange(HS_data_HI, FirstRecord) -->
<!-- nrow(HS_data_HI) -->
<!-- HS_data_HI <-distinct(HS_data_HI, POW.taxonname_HS, .keep_all = TRUE) -->
<!-- nrow(HS_data_HI) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- HS_KB_data_HI <- dplyr::arrange(HS_KB_data_HI, FirstRecord) -->
<!-- nrow(HS_KB_data_HI) -->
<!-- HS_KB_data_HI <-distinct(HS_KB_data_HI, POW.taxonname_HS_KB, .keep_all = TRUE) -->
<!-- nrow(HS_KB_data_HI) -->
<!-- ``` -->
<!-- ### Adding Cultivation and Intro/Nat Date -->

<!-- ```{r} -->
<!-- DATE_CULT <- read.csv(file = "HAPI_data_introdate_cultivation.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F) -->
<!-- DATE_CULT -->
<!-- ``` -->

<!-- ```{r} -->
<!-- tempHS <- merge(HS_data_HI, DATE_CULT, by.x = "POW.taxonname_HS", by.y = "scientificName", all.x = TRUE, all.y = FALSE) -->
<!-- write.csv(tempHS, "tempHS.csv", row.names = FALSE, quote = FALSE) -->
<!-- tempHS -->
<!-- ``` -->

<!-- ```{r} -->
<!-- HS_data_HI <- read.csv(file = "HS_data_fixed.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F) -->
<!-- HS_data_HI -->
<!-- ``` -->


<!-- ```{r} -->
<!-- tempHS_KB <- merge(HS_KB_data_HI, DATE_CULT, by.x = "POW.taxonname_HS_KB", by.y = "scientificName", all.x = TRUE, all.y = FALSE) -->
<!-- write.csv(tempHS_KB, "tempHS_KB.csv", row.names = FALSE, quote = FALSE) -->
<!-- tempHS_KB -->
<!-- ``` -->


<!-- ```{r} -->
<!-- HS_KB_data_HI <- read.csv(file = "HS_KB_data_fixed.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F) -->
<!-- HS_KB_data_HI -->
<!-- ``` -->



<!-- ### Now, let's combine the dataframes -->

<!-- ```{r} -->
<!-- HS_data_HI <- HS_data_HI %>% dplyr::rename(POW.taxonname = POW.taxonname) %>% add_column(Compilation = "Seebens") -->
<!-- HS_KB_data_HI <- HS_KB_data_HI %>% dplyr::rename(POW.taxonname = POW.taxonname)  %>% add_column(Compilation = "Seebens and Brock") -->

<!-- HS_data_HI <- HS_data_HI %>% dplyr::select(POW.taxonname, IPNI_ID, user_supplied_name, Region, Region2, PresentStatus, FirstRecord, FINAL_firsteverdate,	FINAL_firstnatdate, -->
<!--  Source,  DataQuality, Harvest.Point, introductionPathway,	introductionPathway.2, Compilation) -->
<!-- HS_KB_data_HI <- HS_KB_data_HI %>% dplyr::select(POW.taxonname, IPNI_ID, user_supplied_name, Region, Region2, PresentStatus, FirstRecord, FINAL_firsteverdate,	FINAL_firstnatdate, -->
<!--  Source,  DataQuality, Harvest.Point, introductionPathway, introductionPathway.2, Compilation) -->

<!-- ``` -->


<!-- ```{r} -->
<!-- COMBINED <- rbind(HS_data_HI, HS_KB_data_HI) -->
<!-- COMBINED -->
<!-- ``` -->
<!-- ```{r} -->
<!-- write.csv(COMBINED, "COMBINED.csv", row.names = FALSE, quote = FALSE) -->
<!-- ``` -->

```{r}
COMBINED <- read.csv(file = "COMBINED_fixed.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F)
COMBINED
```



## And, let's make trendlines



### For the Seeben's data

```{r}
HS_data_HI <- subset(COMBINED, Compilation_old == "Seebens")
```


### Accumulating the Species
```{r}
alltheyears <- as.data.frame(seq(from = 1000, to = 2020))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(HS_data_HI$FINAL_firsteverdate)) %>%
  dplyr::rename(Year = Var1, Taxa_Naturalized = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Naturalized, sum)
accumulated_A <- cbind(year_count, accumulated_A)
tail(accumulated_A,10)
```


```{r}
alltheyears <- as.data.frame(seq(from = 1000, to = 2020))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(HS_data_HI$FINAL_firstnatdate)) %>%
  dplyr::rename(Year = Var1, Taxa_Naturalized_B = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_B <- accumulate(year_count$Taxa_Naturalized_B, sum)
accumulated_B <- cbind(year_count, accumulated_B)
tail(accumulated_B,10)
```


### Merge 

```{r}
colnames(accumulated_A) <- c("Year", "Taxa_Naturalized", "accumulated")
colnames(accumulated_B) <- c("Year", "Taxa_Naturalized", "accumulated")
accumulated_A$usedate <- "GASFR Data"
accumulated_B$usedate <- "Adjusted GASFR Data"
accumulated_A <- rbind(accumulated_A, accumulated_B)
accumulated_A
```

```{r}
startdate = 1905
enddate = 2005

DATA <- accumulated_A
#DATA$Year <- as.numeric(levels(DATA$Year))
DATA <- subset(DATA, Year >= startdate)
DATA <- subset(DATA, Year <= enddate)
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated), group = usedate, color = usedate)) +
  geom_line(  size=2, alpha=1) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
  guides(colour = guide_legend(reverse=T)) +
 theme_classic() +
  theme(legend.title=element_blank()) +
  labs( title = "", x = "Year", y = "Total Naturalized Species")
ggsave(filename = "HS.png", device = "png", scale = 1, width =6, dpi = 600)
```


### Changepoint Analysis

<!-- ```{r} -->
<!-- DATAfirstever <- subset(DATA, usedate == "Very First Record") -->
<!-- a2 <- changepoint::cpt.meanvar(DATAfirstever$Taxa_Naturalized, penalty = "MBIC", method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 10) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- changepoint::plot(a2, type = "p", ylim = c(0,30)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- c((0 + + DATAfirstever$Year[1]), (changepoint::cpts(a2) + DATAfirstever$Year[1]), DATAfirstever$Year[nrow(DATAfirstever)]) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- changepoint::param.est(a2) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- DATAnatdate <- subset(DATA, usedate != "Very First Record") -->
<!-- a2 <- changepoint::cpt.meanvar(DATAnatdate$Taxa_Naturalized, penalty = "MBIC", method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 10) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- changepoint::plot(a2, type = "p", ylim = c(0,30)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- c((0 +  DATAnatdate$Year[1]), (changepoint::cpts(a2) + DATAnatdate$Year[1]), DATAnatdate$Year[nrow(DATAnatdate)]) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- changepoint::param.est(a2) -->
<!-- ``` -->

### Ridgeline plots to visualize harvest point data


```{r}
DATA2 <- HS_data_HI
DATA2 <- subset(DATA2, FirstRecord >= startdate)
DATA2 <- subset(DATA2, FirstRecord <= enddate)
```

```{r}
  ggplot(DATA2, aes(x = FirstRecord, fill = Harvest.Point)) +
  geom_density(aes(y = ..count..),  stat="binline", bins=50, alpha = 0.65)+
#  scale_x_continuous(breaks = c(1800, 1850, 1900, 1950, 2000)) +
  scale_fill_viridis(discrete = TRUE) +
  theme_classic()

```

### For My Data set

```{r}
HS_KB_data_HI <- subset(COMBINED, Compilation_old == "Seebens and Brock")
HS_KB_data_HI
```




### Accumulating the Species
```{r}
alltheyears <- as.data.frame(seq(from = 1000, to = 2020))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(HS_KB_data_HI$FINAL_firsteverdate)) %>%
  dplyr::rename(Year = Var1, Taxa_Naturalized = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_C <- accumulate(year_count$Taxa_Naturalized, sum)
accumulated_C <- cbind(year_count, accumulated_C)
tail(accumulated_C,10)
```


```{r}
alltheyears <- as.data.frame(seq(from = 1000, to = 2020))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(HS_KB_data_HI$FINAL_firstnatdate)) %>%
  dplyr::rename(Year = Var1, accumulated_D = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_D <- accumulate(year_count$accumulated_D, sum)
accumulated_D <- cbind(year_count, accumulated_D)
tail(accumulated_D,10)
```


### Merge 

```{r}
colnames(accumulated_C) <- c("Year", "Taxa_Naturalized", "accumulated")
colnames(accumulated_D) <- c("Year", "Taxa_Naturalized", "accumulated")
accumulated_C$usedate <- "Very First Record - KB"
accumulated_D$usedate <- "Data from This Publication"
accumulated_C <- rbind(accumulated_C, accumulated_D)
accumulated_C
```

```{r}
#startdate = 1920
#enddate = 2020

DATA <- accumulated_C
#DATA$Year <- as.numeric(levels(DATA$Year))
DATA <- subset(DATA, Year >= startdate)
DATA <- subset(DATA, Year <= enddate)
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated), group = usedate, color = usedate)) +
  geom_line(  size=2, alpha=1) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1800, 1850, 1900, 1950, 2000)) +
  guides(colour = guide_legend(reverse=T)) +
 theme_classic() +
  theme(legend.title=element_blank()) +
  labs( title = "", x = "Year", y = "Total Naturalized Species")
ggsave(filename = "KB.png", device = "png", scale = 1, width =6, dpi = 600)
```


Both analyses on one figure

```{r}
startdate = 1900
enddate = 2020

DATA <- rbind(accumulated_A, accumulated_C)
DATA <- subset(DATA, usedate != "Very First Record - KB")
#DATA$Year <- as.numeric(levels(DATA$Year))
DATA <- subset(DATA, Year >= startdate)
DATA <- subset(DATA, Year <= enddate)
#plotting total species accumulation over time
PanA <- ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated), group = usedate, color = usedate)) +
  geom_line(  size=2, alpha=1) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1900, 1920, 1940,  1960, 1980, 2000, 2020)) +
#  guides(colour = guide_legend(reverse=T)) +
 theme_light() +
  theme( legend.position="top", panel.border = element_blank(),legend.title=element_blank(),
      legend.margin=margin(c(0,0,0,0))) +
  labs( title = "", x = "Year", y = "Total Naturalized Species")
PanA
ggsave(filename = "both.png", device = "png", dpi = 600)


```






### Changepoint Analysis

<!-- ```{r} -->
<!-- a2 <- changepoint::cpt.meanvar(DATA$Taxa_Naturalized, penalty = "MBIC", method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 10) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- changepoint::plot(a2, type = "p", ylim = c(0,30)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- c((0 + + DATA$Year[1]), (changepoint::cpts(a2) + DATA$Year[1]), DATA$Year[nrow(DATA)]) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- changepoint::param.est(a2) -->
<!-- ``` -->


<!-- ### Ridgeline plots to visualize harvest point data -->


<!-- ```{r} -->
<!-- DATA2 <- HS_KB_data_HI -->
<!-- DATA2 <- subset(DATA2, FirstRecord >= startdate) -->
<!-- DATA2 <- subset(DATA2, FirstRecord <= enddate) -->
<!-- ``` -->

<!-- ```{r} -->
<!--   ggplot(DATA2, aes(x = FirstRecord, fill = Harvest.Point)) + -->
<!--   geom_density(aes(y = ..count..),  stat="binline", bins=50, alpha = 0.65)+ -->
<!--   #scale_x_continuous(breaks = c(1800, 1850, 1900, 1950, 2000)) + -->
<!--   scale_fill_viridis(discrete = TRUE) + -->
<!--   theme_classic() -->

<!-- ``` -->

```{r}
COMBINED
```

```{r}

DATA <- COMBINED
#DATA$Year <- as.numeric(levels(DATA$Year))
DATA <- subset(DATA, FINAL_firstnatdate >= startdate)
DATA <- subset(DATA, FINAL_firstnatdate <= enddate)

#DATA <- subset(Per_isl_nat_KtoH_natonly, yearFirst_EVER_Record_statewide >= 1900)

breaks <- c(1900, 1905, 1910,  1915, 1920, 1925, 1930, 1935, 1940, 1945, 1950,  1955, 1960,  1965, 1970,  1975, 1980, 1985, 1990, 1995, 2000, 2005, 2010, 2015, 2020)
labels = c("1905", "1910",  "1915", "1920", "1925", "1930", "1935", "1940", "1945", "1950",  "1955", "1960",  "1965", "1970",  "1975", "1980", "1985", "1990", "1995", "2000", "2005", "2010", "2015", "2020")

DATA$time_binned <- cut(as.numeric(DATA$FINAL_firstnatdate), 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=TRUE, labels = labels)

DATA
```
```{r}
DATA <- DATA %>% distinct(POW.taxonname, time_binned, .keep_all = TRUE)
DATA
```
```{r}
DATA_All <- DATA %>%
  group_by(time_binned, Compilation) %>%
 dplyr::summarize(count = n())
DATA_All

```


 
 
```{r}
 PanB <- DATA_All %>% subset(!is.na(time_binned)) %>% 
#  mutate(Compilation = fct_reorder(Compilation, time_binned)) %>%
ggplot( aes(fill=as.factor(Compilation), y=count, x=time_binned))+ 
    geom_bar(position="fill", stat="identity", size = 0.25,  color = "slategrey") +
    scale_fill_viridis(discrete = T, option = "G") +
  #scale_fill_grey() +
    theme_light() +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 18, b = 0, l = 0)),   plot.margin = unit(c(0.0,0.7,0.3,0.2), "cm"), legend.position="bottom", panel.grid.major = element_blank(),panel.border = element_blank(),legend.title = element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(x = "Time (5-year bins)", y = "Proportion of Known Records")

PanB
ggsave(filename = "temporal_samplingbias.png", device = "png", scale = 1, width =6, height = 3, dpi = 600)
```

```{r}
figure <- multi_panel_figure(
  width = 150, height = 210,
  columns = 1, rows = 2)

  figure %>% fill_panel(PanA, row = 1, column = 1) %>%
  fill_panel(PanB, row = 2, column = 1) %>%
  
  save_multi_panel_figure(filename = "Figure5_simstudy.png", dpi = 600)
```


## Model Fitting

```{r}
startdate = 1950
enddate = 2005


DATA <- rbind(accumulated_A, accumulated_C)
DATA <- subset(DATA, usedate != "Very First Record - KB")
#DATA$Year <- as.numeric(levels(DATA$Year))
DATA <- subset(DATA, Year >= startdate)
DATA <- subset(DATA, Year <= enddate)
DATA
```

```{r}
unique(DATA$usedate)
```
```{r}

#DATA <- subset(Per_isl_nat_KtoH_natonly, yearFirst_EVER_Record_statewide >= 1900)

breaks <- seq(from = 1800, to = 2020, by = 5)
labels = c("1805", "1810", "1815", "1820", "1825", "1830", "1835", "1840", "1845", "1850", "1855", "1860", "1865", "1870", "1875", "1880", "1885", "1890", "1895", "1900", "1905", "1910",  "1915", "1920", "1925", "1930", "1935", "1940", "1945", "1950",  "1955", "1960",  "1965", "1970",  "1975", "1980", "1985", "1990", "1995", "2000", "2005", "2010", "2015", "2020")

DATA$time_binned <- cut(as.numeric(DATA$Year), 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=TRUE, labels = labels)

DATA
```

```{r}
DATA_1 <- DATA %>%
  group_by(time_binned, usedate) %>%
 dplyr::summarize(Taxa_Naturalized = sum(Taxa_Naturalized))
DATA_1
```


```{r}
DATA_GASFR <- subset(DATA_1, usedate == "Data from This Publication")
DATA_GASFR <- subset(DATA_GASFR, !is.na(Taxa_Naturalized))
DATA_GASFR <- subset(DATA_GASFR, !is.na(time_binned))
DATA_GASFR$time_binned <- as.numeric(DATA_GASFR$time_binned)

##Load data
x <- DATA_GASFR$time_binned
y <- DATA_GASFR$Taxa_Naturalized

plot(x, y)

```
```{r}
all(is.na(x))
all(is.na(y))
```
```{r}
unique(x)
```



```{r}
# set the margins
tmpmar <- par("mar")
tmpmar[3] <- 0.5
par(mar=tmpmar)

# get underlying plot
x <- DATA_GASFR$time_binned
y <- DATA_GASFR$Taxa_Naturalized
plot(x, y, pch=20)

# basic straight line of fit
fit.lm.o <- glm(y~x)
co.lm.o <- coef(fit.lm.o)
co.lm.o
abline(fit.lm.o, col="blue", lwd=2)

# exponential
#fm_lm <- lm(log(y) ~ x, DATA_GASFR)
#st <- list(a = exp(coef(fm_lm)[1]), b = coef(fm_lm)[2])
#f.ex.o <- function(x,a,b) {a * exp(b * x)}
#fit.ex.o <- nls(y ~ f.ex.o(x,a,b),
#              start = st, control = list(maxiter = 1000), trace = T)  
#co.ex.o <- coef(fit.ex.o)
#co.ex.o
#curve(f.ex.o(x, a=co.ex.o[1], b=co.ex.o[2]), add = TRUE, col="green", lwd=2) 

# logarithmic
f.ln.o <- function(x,a,b) {a * log(x) + b}
fit.ln.o <- nls(y ~ f.ln.o(x,a,b), start = c(a=1, b=1)) 
co.ln.o <- coef(fit.ln.o)
curve(f.ln.o(x, a=co.ln.o[1], b=co.ln.o[2]), add = TRUE, col="orange", lwd=2) 

# polynomial
f.po.o <- function(x,a,b,d) {(a*x^2) + (b*x) + d}
fit.po.o <- nls(y ~ f.po.o(x,a,b,d), start = c(a=1, b=1, d=1)) 
co.po.o <- coef(fit.po.o)
curve(f.po.o(x, a=co.po.o[1], b=co.po.o[2], d=co.po.o[3]), add = TRUE, col="pink", lwd=2)

# legend
legend("topleft",
    legend=c("linear","exponential","logarithmic","polynomial"),
    col=c("blue","green","orange","pink"),
    lwd=2,
    )

```

```{r}
AIC(fit.lm.o)
#AIC(fit.ex.o)
AIC(fit.ln.o)
AIC(fit.po.o)
```

```{r}
model_w1 <- drm(y ~ x, fct = W1.4()) #Weibull type 1 model
model_w2 <- drm(y ~ x, fct = W2.4()) #Weibull type 2 model
plot(model_w1,  col="yellow", lwd=2.5)
plot(model_w2, add=T, col="orange", lwd=2.5)
```


```{r}
#model_line <- nls(y ~ NLS.linear(x, a, b)) #Linear model
model_line <- drm(y ~ x, fct = DRC.linear())

model_poly <- drm(y ~ x, fct = DRC.poly2()) #2nd order polynomial model
model_exp <- drm(y ~ x, fct = DRC.expoGrowth()) #Exponential growth model
model_decay <- drm(y ~ x, fct = DRC.expoDecay()) #Exponential decay model
```


```{r}
model_asym <- drm(y ~ x, fct = DRC.asymReg()) #Asymptotic model
model_pwr <- drm(y ~ x, fct = DRC.powerCurve()) #Power law model
model_logth <- drm(y ~ x, fct = DRC.logCurve()) #Logarithmic model
```


```{r}
#model_mm <- drm(y ~ x, fct = MM.2()) #Michaelis-Menten logarithmic model: rapid growth and then plateau
model_sig <- drm(y ~ x, fct = L.3()) #Sigmoidal/logistic model
```


```{r}
#model_g1 <- drm(y ~ x, fct = G.3()) #Gompertz sigmoidal curve with long beginning lag
#model_g2 <- drm(y ~ x, fct = E.3()) #Gompertz sigmoidal curve without beginning lag
```


```{r}
model_ll <- drm(y ~ x, fct = LL.4()) #Log-logistic model
model_w1 <- drm(y ~ x, fct = W1.4()) #Weibull type 1 model
model_w2 <- drm(y ~ x, fct = W2.4()) #Weibull type 2 model
```
```{r}
mod.bragg <- drm(y ~ x, fct = DRC.bragg.3() )
mod.bragg2 <- drm(y ~ x, fct = DRC.bragg.4() )

```

```{r}
##Take a look
#Polynomial curve
#plot(model_poly,col="darkgrey", lwd=2.5)



#Convex/concave models
plot(model_decay, col="purple",lwd=2.5)
plot(model_asym, add=T, col="darkmagenta", lwd=2.5)
plot(model_exp, add=T,col="blue", lwd=2.5)
plot(model_pwr, add=T,col="darkgrey", lwd=2.5)
plot(model_logth, add=T,col="blue", lwd=2.5)
#plot(model_mm, add=T,col="green", lwd=2.5)

#Sigmoidal curves
plot(model_sig, add=T,col="red", lwd=2.5)
#plot(model_g1, add=T,col="red", lwd=2.5)
#plot(model_g2, add=T,col="pink", lwd=2.5)
plot(model_ll, add=T,col="darkorange", lwd=2.5)
plot(model_w1, add=T, col="yellow", lwd=2.5)
plot(model_w2, add=T, col="orange", lwd=2.5)

# hump-shaped curves
plot(mod.bragg, add = T, col = "darkgreen",  lwd=2.5) 
plot(mod.bragg2, add = T, col = "darkgreen")

#Linear
plot(model_line, add = T, col="black",  lwd=2.5)

```

```{r}
##Evaluate fit
AIC(model_line)
AIC(model_poly)
AIC(model_exp)
AIC(model_decay)
AIC(model_asym)
AIC(model_pwr)
AIC(model_logth)
AIC(model_sig)
AIC(model_ll)
AIC(model_w1)
AIC(model_w2)
AIC(mod.bragg)
AIC(mod.bragg2)
```







## Let's compare trends for THE WHOLE PACIFIC ISLANDS

### Subsetting both datasets to Hawaii Only

```{r}
HS_data_PAC <- HS_data
nrow(HS_data_PAC)
HS_KB_data_PAC <- HS_KB_data
nrow(HS_KB_data_PAC)
```
### Make sure there are no duplicates (if there are, we should take only the first record)

```{r}
HS_data_PAC <- dplyr::arrange(HS_data_PAC, FirstRecord)
nrow(HS_data_PAC)
HS_data_PAC <-distinct(HS_data_PAC, POW.taxonname_HS, .keep_all = TRUE)
nrow(HS_data_PAC)
```
```{r}
HS_KB_data_PAC <- dplyr::arrange(HS_KB_data_PAC, FirstRecord)
nrow(HS_KB_data_PAC)
HS_KB_data_PAC <-distinct(HS_KB_data_PAC, POW.taxonname_HS_KB, .keep_all = TRUE)
nrow(HS_KB_data_PAC)
```

## And, let's make trendlines



### For the Seeben's data

```{r}
HS_data_PAC
```
```{r}
str(HS_data_PAC)
```


### Accumulating
```{r}
alltheyears <- as.data.frame(seq(from = 1000, to = 2020))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(HS_data_PAC$FirstRecord)) %>%
  dplyr::rename(Year = Var1, Taxa_Naturalized = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Naturalized, sum)
accumulated_A <- cbind(year_count, accumulated_A)
tail(accumulated_A,10)
```


```{r}
DATA <- accumulated_A
#DATA$Year <- as.numeric(levels(DATA$Year))
DATA <- subset(DATA, Year >= 1810)
DATA <- subset(DATA, Year <= 2010)
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_A))) +
  geom_line( color = "black", size=2, alpha=1) +
 # scale_color_viridis(discrete = TRUE) +
#  scale_x_continuous(breaks = c(1800, 1850, 1900, 1950, 2000)) +
  guides(colour = guide_legend(reverse=T)) +
 theme_classic() +
  theme(legend.title=element_blank()) +
  labs( title = "HS Pacific", x = "Year", y = "Total Naturalized Species")
ggsave(filename = "HS_PAC.png", device = "png", scale = 1, width =6, dpi = 600)
```






### Changepoint Analysis

```{r}
a2 <- changepoint::cpt.meanvar(DATA$Taxa_Naturalized, penalty = "MBIC", method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 10)
```

```{r}
changepoint::plot(a2, type = "p", ylim = c(0,30))
```

```{r}
c((0 + + DATA$Year[1]), (changepoint::cpts(a2) + DATA$Year[1]), DATA$Year[nrow(DATA)])
```

```{r}
changepoint::param.est(a2)
```


### For My Data set

```{r}
HS_KB_data_PAC
```
```{r}
str(HS_KB_data_PAC)
```


### Accumulating
```{r}
alltheyears <- as.data.frame(seq(from = 1000, to = 2020))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(HS_KB_data_PAC$FirstRecord)) %>%
  dplyr::rename(Year = Var1, Taxa_Naturalized = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Naturalized, sum)
accumulated_A <- cbind(year_count, accumulated_A)
tail(accumulated_A,10)
```


```{r}
DATA <- accumulated_A
#DATA$Year <- as.numeric(levels(DATA$Year))
DATA <- subset(DATA, Year >= 1810)
DATA <- subset(DATA, Year <= 2010)
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_A))) +
  geom_line( color = "black", size=2, alpha=1) +
 # scale_color_viridis(discrete = TRUE) +
#  scale_x_continuous(breaks = c(1800, 1850, 1900, 1950, 2000)) +
  guides(colour = guide_legend(reverse=T)) +
 theme_classic() +
  theme(legend.title=element_blank()) +
  labs( title = "KB Pacific", x = "Year", y = "Total Naturalized Species")
ggsave(filename = "KB_PAC.png", device = "png", scale = 1, width =6, dpi = 600)
```

### Changepoint Analysis

```{r}
a2 <- changepoint::cpt.meanvar(DATA$Taxa_Naturalized, penalty = "MBIC", method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)
```

```{r}
changepoint::plot(a2, type = "p", ylim = c(0,30))
```

```{r}
c((0 + + DATA$Year[1]), (changepoint::cpts(a2) + DATA$Year[1]), DATA$Year[nrow(DATA)])

changepoint::param.est(a2)
```
