---
title: "simstudy_PART3_harvestpoints"
author: "Kelsey"
date: "10/2/2021"
output: html_document
---

```{r, }
require(knitr)
```
```{r}
#knitr::opts_knit$set(root.dir = "C:/Users/Kelsey/Documents/codingwork/simulation_study") #set the working directory
```
```{r}
#get the needed packages
if(!require("pacman")){
	install.packages("pacman")
	library(pacman)}
p_load("dplyr", "ggplot2", "viridis", "tidyr",  "fGarch", "sn", "purrr", "forcats", "stats", "changepoint", "changepoint.np")
```

```{r}
sim_spp <- read.csv(file = "sim_spp.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F)
accumulated <- read.csv(file = "accumulated.csv",  header=T, sep=',', na.strings=c("","NA"), stringsAsFactors=F)

accumulated
```

```{r}
#accumulated$Taxa_Naturalized_centered <- accumulated$Taxa_Naturalized - mean(accumulated$Taxa_Naturalized)
plot(Taxa_Naturalized ~  Year, data = accumulated)
```
```{r}
#accumulated$Taxa_Naturalized_centered <- accumulated$Taxa_Naturalized - mean(accumulated$Taxa_Naturalized)
hist(sim_spp$Naturalization_Year)
```

### Changepoint Analysis

```{r}
a2 <- changepoint::cpt.meanvar(accumulated$Taxa_Naturalized, penalty = "MBIC", method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)
```

```{r}
changepoint::plot(a2)
```
```{r}
changepoint::param.est(a2)
```
```{r}
accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
```



# ACCESSIBILITY PART

# Scenario 5: Harvest Points Introduce different cutoff points : ID'd vouchers


## Simulate Hypothetical Lag Times

For this scenario, let's say that all lags have a mean of 10 years

```{r}
sim_spp5 <- sim_spp
```

##### 3.0
Simulate the lag between introduction and naturalization

```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Intro_Nat_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
Intro_Nat_Lag <- subset(Intro_Nat_Lag, Intro_Nat_Lag >=0 & Intro_Nat_Lag <=20)
Intro_Nat_Lag <- as.data.frame(sample(Intro_Nat_Lag$Intro_Nat_Lag, 1000))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"

# #plotting histogram
# ggplot(Intro_Nat_Lag, aes(x = Intro_Nat_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Intro_Nat_Lag)
# #Nat_Detect_Lag
```

##### 3.1
Then, simulate the lag between naturalization and detection
let's say this takes around 5 years, with some taking a very long time
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Nat_Detect_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
Nat_Detect_Lag <- subset(Nat_Detect_Lag, Nat_Detect_Lag >=0 & Nat_Detect_Lag <=20)
Nat_Detect_Lag <- as.data.frame(sample(Nat_Detect_Lag$Nat_Detect_Lag, 1000))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
#plotting histogram
# ggplot(Nat_Detect_Lag, aes(x = Nat_Detect_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Nat_Detect_Lag)
#Nat_Detect_Lag
```
##### 3.2
Second, simulate the lag between detection and evidence collection
let's say this takes around 1 years, with some taking longer, up to 5 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Detect_Vouch_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Detect_Vouch_Lag) <- "Detect_Vouch_Lag"
Detect_Vouch_Lag <- subset(Detect_Vouch_Lag, Detect_Vouch_Lag >=0 & Detect_Vouch_Lag <=20)
Detect_Vouch_Lag <- as.data.frame(sample(Detect_Vouch_Lag$Detect_Vouch_Lag, 1000))
colnames(Detect_Vouch_Lag) <- "Detect_Vouch_Lag"
#plotting histogram
# ggplot(Detect_Vouch_Lag, aes(x = Detect_Vouch_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Detect_Vouch_Lag)

```

##### 3.3

third, simulate the lag between vouchering and identification
let's say this takes around 1 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Vouch_ID_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
Vouch_ID_Lag <- subset(Vouch_ID_Lag, Vouch_ID_Lag >=0 & Vouch_ID_Lag <=20)
Vouch_ID_Lag <- as.data.frame(sample(Vouch_ID_Lag$Vouch_ID_Lag, 1000))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
#plotting histogram
# ggplot(Vouch_ID_Lag, aes(x = Vouch_ID_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Vouch_ID_Lag)
```
##### 3.4
Then, simulate the lag between IDing and reporting
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

ID_Report_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
ID_Report_Lag <- subset(ID_Report_Lag, ID_Report_Lag >=0 & ID_Report_Lag <=20)
ID_Report_Lag <- as.data.frame(sample(ID_Report_Lag$ID_Report_Lag, 1000))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
# #plotting histogram
# ggplot(ID_Report_Lag, aes(x = ID_Report_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(ID_Report_Lag)
```


HMMM!  I'm not sure this really works like this.  Isn't it more like someone comes in, and then just compiles everything that they can at once, no matter if they were reported 1 year ago, or 30?  Yes....

##### 3.5
Then, simulate the lag between reporting  and when someone makes a resource that is compiled and formatted
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Report_Compile_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
Report_Compile_Lag <- subset(Report_Compile_Lag, Report_Compile_Lag >=0 & Report_Compile_Lag <=20)
Report_Compile_Lag <- as.data.frame(sample(Report_Compile_Lag$Report_Compile_Lag, 1000))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
#plotting histogram
# ggplot(Report_Compile_Lag, aes(x = Report_Compile_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Report_Compile_Lag)
```

So, it may be arbitrary, but the report would compile all species up to that date.

##### 3.6 Okay, lets add up the lags
So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag,

All created by adding or subtracting from the naturalization year
```{r}
sim_spp5$Intro_Year <- sim_spp5$Naturalization_Year - Intro_Nat_Lag$Intro_Nat_Lag
sim_spp5$Detect_Year <- sim_spp5$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag
sim_spp5$Voucher_Year <- sim_spp5$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag
sim_spp5$ID_Year <- sim_spp5$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag
sim_spp5$Report_Year <- sim_spp5$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag


sim_spp5$Compile_Year <- sim_spp5$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag + Report_Compile_Lag$Report_Compile_Lag


sim_spp5$All_Processing_Lags <- Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag
#reordering according to Report Year
sim_spp5 <- sim_spp5 %>% dplyr::select(Species_IDNo, Intro_Year, Naturalization_Year, Detect_Year, Voucher_Year, ID_Year, Report_Year, Compile_Year, All_Processing_Lags)
sim_spp5
```

### Cropping the Dataset because we can't know about species that haven't been reported yet

<!-- What's the lags look like for all species -->
<!-- ```{r} -->
<!-- #plotting histogram -->
<!-- ggplot(sim_spp5, aes(x = All_Processing_Lags)) + -->
<!--   geom_density(fill="#440154FF", color="#e9ecef", alpha=0.6) + -->
<!--  # scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) + -->
<!--   theme_classic() -->
<!-- ``` -->

MAJOR DIFFERENCE: you get to crop based on the ID year, because you're looking at the vouchers themselves
How many species did we lose that hadn't been reported yet?
```{r}
sim_spp5_crop <- subset(sim_spp5, ID_Year <= 2020)
```

```{r}
nrow(sim_spp5) -nrow(sim_spp5_crop)
```
```{r}
retrieval_rate = 1
sampled <- as.data.frame(sample(sim_spp5_crop$Species_IDNo, (nrow(sim_spp5_crop)* retrieval_rate)))
colnames(sampled) <- "Species_IDNo"
sim_spp5_crop_sample <- merge(sampled, sim_spp5_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE)
sim_spp5_crop_sample
```

## Step 4. Analyse Trends - A) Voucher Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(sim_spp5_crop_sample$Voucher_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Vouchered = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum)
accumulated_A <- cbind(year_count, accumulated_A)
#accumulated_A
```

## Step 5. Analyse Trends - B) Report Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp5_crop_sample$Report_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Reported = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_B <- accumulate(year_count$Taxa_Reported, sum)
accumulated_B <- cbind(year_count, accumulated_B)
#accumulated_B
```
## Step 6. Analyse Trends - C) Intro Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp5_crop_sample$Intro_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Introduced = Freq)
year_count

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_C <- accumulate(year_count$Taxa_Introduced, sum)
accumulated_C <- cbind(year_count, accumulated_C)
#accumulated_C
```

## Step 6. Analyse Trends - D) Compiled Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp5_crop_sample$Compile_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Compiled = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_D <- accumulate(year_count$Taxa_Compiled, sum)
accumulated_D <- cbind(year_count, accumulated_D)
#accumulated_D
```

#### Changepoint - Evidence Collection

```{r}
plot(accumulated_A$Taxa_Vouchered)
```
```{r}
a1 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

plot(a1, diagnostic = TRUE)
```
```{r}
selectedcp = 3
#changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)")
changepoint::plot(a1, ncpts = selectedcp)
```
```{r}
changepoint::cpts.full(a1) + (accumulated_A$Year[1] -1)
```
```{r}
firstrow = 4
secondrow = firstrow +1
a2 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "Manual",  pen.value=((pen.value.full(a1)[firstrow] + pen.value.full(a1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A))
chpoints
chpoints + accumulated_A$Year[1]
segmeans <- changepoint::param.est(a2)$mean
segmeans
segvars <- (changepoint::param.est(a2)$variance)
sqrt(segvars)
```
```{r}
seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1])
seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2])
seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3])
seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4])
seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5])
seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6])
seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7])
seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8])
seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1])
seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2])
seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3])
seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4])
seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5])
seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6])
seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7])
seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8])

segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean)
segwmeans <- segwmeans[!is.na(segwmeans)]
segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar)
segwvars <- segwvars[!is.na(segwvars)]
```
```{r}
startptindex = 3
endptindex = 4

print("How many species did we lose that hadn't been reported yet?")
spplostharvest <- nrow(sim_spp5) -nrow(sim_spp5_crop)
spplostharvest

spplostretrieval <- nrow(sim_spp5_crop) - nrow(sim_spp5_crop_sample)
spplostretrieval

print("start seg length, mean, std: ")
chpoints[startptindex]
round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2)
round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2)

print("truncated seg length, mean, std:")
chpoints[endptindex] - chpoints[startptindex]
round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2)
round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2)

print("ending seg length, mean, std:")
nrow(accumulated_A) - chpoints[endptindex]
round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2)
round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2)
```
```{r}
changepoint::plot(a2, type='p')
```



#### Changepoints - Intro Date

```{r}
plot(accumulated_C$Taxa_Introduced)
```
```{r}
c1 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

plot(c1, diagnostic = TRUE)
```
```{r}
selectedcp = 3
#changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)")
changepoint::plot(c1, ncpts = selectedcp)
```

```{r}
changepoint::cpts.full(c1) + as.numeric(accumulated_C$Year[1]) -1
```
```{r}
firstrow = 4
secondrow = firstrow +1
c2 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "Manual",  pen.value=((pen.value.full(c1)[firstrow] + pen.value.full(c1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5)

chpoints <- c( 0, changepoint::cpts(c2), nrow(accumulated_C))
chpoints
chpoints + as.numeric(accumulated_C$Year[1])
segmeans <- changepoint::param.est(c2)$mean
segmeans
segvars <- (changepoint::param.est(c2)$variance)
sqrt(segvars)
```
```{r}
seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1])
seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2])
seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3])
seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4])
seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5])
seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6])
seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7])
seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8])
seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1])
seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2])
seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3])
seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4])
seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5])
seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6])
seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7])
seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8])

segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean)
segwmeans <- segwmeans[!is.na(segwmeans)]
segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar)
segwvars <- segwvars[!is.na(segwvars)]
```
```{r}
startptindex = 2
endptindex = 3

print("How many species did we lose that hadn't been reported yet?")
spplostharvest <- nrow(sim_spp5) -nrow(sim_spp5_crop)
spplostharvest

spplostretrieval <- nrow(sim_spp5_crop) - nrow(sim_spp5_crop_sample)
spplostretrieval

print("start seg length, mean, std: ")
chpoints[startptindex]
round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2)
round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2)

print("truncated seg length, mean, std:")
chpoints[endptindex] - chpoints[startptindex]
round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2)
round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2)

print("ending seg length, mean, std:")
nrow(accumulated_A) - chpoints[endptindex]
round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2)
round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2)
```
```{r}
changepoint::plot(c2, type='p')
```



```{r}
#making into a single dataframe
#accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered)
accumulated_B  <- accumulated_B %>% dplyr::rename(accumulated_species = accumulated_B, No_Taxa = Taxa_Reported)
accumulated_C  <- accumulated_C %>% dplyr::rename(accumulated_species = accumulated_C, No_Taxa = Taxa_Introduced)
accumulated_D  <- accumulated_D %>% dplyr::rename(accumulated_species = accumulated_D, No_Taxa = Taxa_Compiled)
accumulated$Analysis_Type <- "Hypothetical Naturalization Date"
accumulated_A$Analysis_Type <- "Evidence Collection Date"
accumulated_B$Analysis_Type <- "Report Date"
accumulated_C$Analysis_Type <- "Introduction Date"
accumulated_D$Analysis_Type <- "Compiled Date"

accumulated_A_B_C_D <- rbind(accumulated, accumulated_A, accumulated_B, accumulated_C, accumulated_D)
accumulated_A_B_C_D
```

```{r}
# DATA <- accumulated_A_B_C_D %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
# DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
# #plotting total species accumulation over time
# ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(No_Taxa), group = Analysis_Type, color = Analysis_Type)) +
#   geom_smooth( size=1.5, alpha=0.8, level = 0, method = "loess", fullrange=TRUE) +
#   scale_color_viridis(discrete = TRUE) +
#   scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
#   guides(colour = guide_legend(reverse=T)) +
#  theme_classic() +
#   labs(title = "Scenario5 Nat Rate", x = "Year", y = "Number of Species")
# ggsave(filename = "Scenario5_ rate.png", device = "png", scale = 1, width =6, dpi = 600)
```

```{r}
DATA <- accumulated_A_B_C_D# %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
DATA <- subset(DATA, DATA$Analysis_Type != "Report Date")
DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) +
  geom_line( size=1.5, alpha=0.70) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
  guides(colour = guide_legend(reverse=T)) +
 theme_classic() +
  theme(legend.title=element_blank()) +
  labs(title = "Harvest Point = Identified Specimens deposited in Natural History Collections", x = "Year", y = "Number of Species")
ggsave(filename = "HP_ided_specimens.png", device = "png", scale = 1, width =6, dpi = 600)
```

# Scenario 6: Harvest Points Introduce different cutoff points : REPORTS

- prevents leveling off, but excludes species because not all vouchers mentions their status


## Simulate Hypothetical Lag Times

For this scenario, let's say that all lags have a mean of 10 years

```{r}
sim_spp6 <- sim_spp
```

##### 3.0
Simulate the lag between introduction and naturalization

```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Intro_Nat_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
Intro_Nat_Lag <- subset(Intro_Nat_Lag, Intro_Nat_Lag >=0 & Intro_Nat_Lag <=20)
Intro_Nat_Lag <- as.data.frame(sample(Intro_Nat_Lag$Intro_Nat_Lag, 1000))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"

#plotting histogram
# ggplot(Intro_Nat_Lag, aes(x = Intro_Nat_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Intro_Nat_Lag)
#Nat_Detect_Lag
```

##### 3.1
Then, simulate the lag between naturalization and detection
let's say this takes around 5 years, with some taking a very long time
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Nat_Detect_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
Nat_Detect_Lag <- subset(Nat_Detect_Lag, Nat_Detect_Lag >=0 & Nat_Detect_Lag <=20)
Nat_Detect_Lag <- as.data.frame(sample(Nat_Detect_Lag$Nat_Detect_Lag, 1000))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
#plotting histogram
# ggplot(Nat_Detect_Lag, aes(x = Nat_Detect_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Nat_Detect_Lag)
#Nat_Detect_Lag
```
##### 3.2
Second, simulate the lag between detection and evidence collection
let's say this takes around 1 years, with some taking longer, up to 5 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Detect_Vouch_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Detect_Vouch_Lag) <- "Detect_Vouch_Lag"
Detect_Vouch_Lag <- subset(Detect_Vouch_Lag, Detect_Vouch_Lag >=0 & Detect_Vouch_Lag <=20)
Detect_Vouch_Lag <- as.data.frame(sample(Detect_Vouch_Lag$Detect_Vouch_Lag, 1000))
colnames(Detect_Vouch_Lag) <- "Detect_Vouch_Lag"
#plotting histogram
# ggplot(Detect_Vouch_Lag, aes(x = Detect_Vouch_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Detect_Vouch_Lag)

```

##### 3.3

third, simulate the lag between vouchering and identification
let's say this takes around 1 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Vouch_ID_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
Vouch_ID_Lag <- subset(Vouch_ID_Lag, Vouch_ID_Lag >=0 & Vouch_ID_Lag <=20)
Vouch_ID_Lag <- as.data.frame(sample(Vouch_ID_Lag$Vouch_ID_Lag, 1000))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
#plotting histogram
# ggplot(Vouch_ID_Lag, aes(x = Vouch_ID_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Vouch_ID_Lag)
```
##### 3.4
Then, simulate the lag between IDing and reporting
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

ID_Report_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
ID_Report_Lag <- subset(ID_Report_Lag, ID_Report_Lag >=0 & ID_Report_Lag <=20)
ID_Report_Lag <- as.data.frame(sample(ID_Report_Lag$ID_Report_Lag, 1000))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
#plotting histogram
# ggplot(ID_Report_Lag, aes(x = ID_Report_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(ID_Report_Lag)
```


##### 3.5
Then, simulate the lag between reporting  and when someone makes a resource that is compiled and formatted
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Report_Compile_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
Report_Compile_Lag <- subset(Report_Compile_Lag, Report_Compile_Lag >=0 & Report_Compile_Lag <=20)
Report_Compile_Lag <- as.data.frame(sample(Report_Compile_Lag$Report_Compile_Lag, 1000))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
#plotting histogram
# ggplot(Report_Compile_Lag, aes(x = Report_Compile_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Report_Compile_Lag)
```

So, it may be arbitrary, but the report would compile all species up to that date.

##### 3.6 Okay, lets add up the lags
So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag,

All created by adding or subtracting from the naturalization year
```{r}
sim_spp6$Intro_Year <- sim_spp6$Naturalization_Year - Intro_Nat_Lag$Intro_Nat_Lag
sim_spp6$Detect_Year <- sim_spp6$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag
sim_spp6$Voucher_Year <- sim_spp6$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag
sim_spp6$ID_Year <- sim_spp6$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag
sim_spp6$Report_Year <- sim_spp6$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag


sim_spp6$Compile_Year <- sim_spp6$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag + Report_Compile_Lag$Report_Compile_Lag


sim_spp6$All_Processing_Lags <- Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag
#reordering according to Report Year
sim_spp6 <- sim_spp6 %>% dplyr::select(Species_IDNo, Intro_Year, Naturalization_Year, Detect_Year, Voucher_Year, ID_Year, Report_Year, Compile_Year, All_Processing_Lags)
#sim_spp6
```

### Cropping the Dataset because we can't know about species that haven't been reported yet

What's the lags look like for all species
```{r}
# #plotting histogram
# ggplot(sim_spp6, aes(x = All_Processing_Lags)) +
#   geom_density(fill="#440154FF", color="#e9ecef", alpha=0.6) +
#  # scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
#   theme_classic()
```

MAJOR DIFFERENCE: you get to crop based on the ID year, because you're looking at the vouchers themselves
How many species did we lose that hadn't been reported yet?
```{r}
sim_spp6_crop <- subset(sim_spp6, Report_Year <= 2020)
sim_spp6_crop
```

```{r}
nrow(sim_spp6) -nrow(sim_spp6_crop)
```
```{r}
retrieval_rate = 1
sampled <- as.data.frame(sample(sim_spp6_crop$Species_IDNo, (nrow(sim_spp6_crop)* retrieval_rate)))
colnames(sampled) <- "Species_IDNo"
sim_spp6_crop_sample <- merge(sampled, sim_spp6_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE)
sim_spp6_crop_sample
```


## Step 4. Analyse Trends - A) Voucher Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(sim_spp6_crop_sample$Voucher_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Vouchered = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum)
accumulated_A <- cbind(year_count, accumulated_A)
accumulated_A
```

## Step 5. Analyse Trends - B) Report Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp6_crop_sample$Report_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Reported = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_B <- accumulate(year_count$Taxa_Reported, sum)
accumulated_B <- cbind(year_count, accumulated_B)
accumulated_B
```
## Step 6. Analyse Trends - C) Intro Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp6_crop_sample$Intro_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Introduced = Freq)
year_count

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_C <- accumulate(year_count$Taxa_Introduced, sum)
accumulated_C <- cbind(year_count, accumulated_C)
accumulated_C
```


## Step 6. Analyse Trends - D) Compiled Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp6_crop_sample$Compile_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Compiled = Freq)
year_count

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_D <- accumulate(year_count$Taxa_Compiled, sum)
accumulated_D <- cbind(year_count, accumulated_D)
accumulated_D
```


```{r}
#making into a single dataframe
#accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered)
accumulated_B  <- accumulated_B %>% dplyr::rename(accumulated_species = accumulated_B, No_Taxa = Taxa_Reported)
accumulated_C  <- accumulated_C %>% dplyr::rename(accumulated_species = accumulated_C, No_Taxa = Taxa_Introduced)
accumulated_D  <- accumulated_D %>% dplyr::rename(accumulated_species = accumulated_D, No_Taxa = Taxa_Compiled)
accumulated$Analysis_Type <- "Hypothetical Naturalization Date"
accumulated_A$Analysis_Type <- "Evidence Collection Date"
accumulated_B$Analysis_Type <- "Report Date"
accumulated_C$Analysis_Type <- "Introduction Date"
accumulated_D$Analysis_Type <- "Compiled Date"

accumulated_A_B_C_D <- rbind(accumulated, accumulated_A, accumulated_B, accumulated_C, accumulated_D)
accumulated_A_B_C_D
```

# ```{r}
# DATA <- accumulated_A_B_C_D %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
# DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
# #plotting total species accumulation over time
# ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(No_Taxa), group = Analysis_Type, color = Analysis_Type)) +
#   geom_smooth( size=1.5, alpha=0.8, level = 0, method = "loess", fullrange=TRUE) +
#   scale_color_viridis(discrete = TRUE) +
#   scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
#   guides(colour = guide_legend(reverse=T)) +
#  theme_classic() +
#   labs(title = "Scenario6 Nat Rate", x = "Year", y = "Number of Species")
# ggsave(filename = "Scenario6_ rate.png", device = "png", scale = 1, width =6, dpi = 600)
# ```

```{r}
DATA <- accumulated_A_B_C_D# %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
DATA <- subset(DATA, DATA$Analysis_Type != "Report Date")
DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) +
  geom_line( size=1.5, alpha=0.70) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
  guides(colour = guide_legend(reverse=T)) +
 theme_classic() +
  theme(legend.title=element_blank()) +
  labs(title = "Harvest Point = Published Reports on Naturalization Status", x = "Year", y = "Number of Species")
ggsave(filename = "Scenario6.png", device = "png", scale = 1, width =6, dpi = 600)
```

# Scenario 7: Access Points Introduce different cutoff points : CONTINUALLY UPDATED DATABASE WITH LAG BETWEEN REPORT AND COMPILE

- prevents leveling off, but excludes species because not all vouchers mentions their status


## Simulate Hypothetical Lag Times

For this scenario, let's say that all lags have a mean of 10 years

```{r}
sim_spp7 <- sim_spp
```

##### 3.0
Simulate the lag between introduction and naturalization

```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Intro_Nat_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
Intro_Nat_Lag <- subset(Intro_Nat_Lag, Intro_Nat_Lag >=0 & Intro_Nat_Lag <=20)
Intro_Nat_Lag <- as.data.frame(sample(Intro_Nat_Lag$Intro_Nat_Lag, 1000))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"

#plotting histogram
# ggplot(Intro_Nat_Lag, aes(x = Intro_Nat_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Intro_Nat_Lag)
#Nat_Detect_Lag
```

##### 3.1
Then, simulate the lag between naturalization and detection
let's say this takes around 5 years, with some taking a very long time
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Nat_Detect_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
Nat_Detect_Lag <- subset(Nat_Detect_Lag, Nat_Detect_Lag >=0 & Nat_Detect_Lag <=20)
Nat_Detect_Lag <- as.data.frame(sample(Nat_Detect_Lag$Nat_Detect_Lag, 1000))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
#plotting histogram
# ggplot(Nat_Detect_Lag, aes(x = Nat_Detect_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Nat_Detect_Lag)
# #Nat_Detect_Lag
```
##### 3.2
Second, simulate the lag between detection and evidence collection
let's say this takes around 1 years, with some taking longer, up to 5 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Detect_Vouch_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Detect_Vouch_Lag) <- "Detect_Vouch_Lag"
Detect_Vouch_Lag <- subset(Detect_Vouch_Lag, Detect_Vouch_Lag >=0 & Detect_Vouch_Lag <=20)
Detect_Vouch_Lag <- as.data.frame(sample(Detect_Vouch_Lag$Detect_Vouch_Lag, 1000))
colnames(Detect_Vouch_Lag) <- "Detect_Vouch_Lag"
#plotting histogram
# ggplot(Detect_Vouch_Lag, aes(x = Detect_Vouch_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Detect_Vouch_Lag)

```

##### 3.3

third, simulate the lag between vouchering and identification
let's say this takes around 1 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Vouch_ID_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
Vouch_ID_Lag <- subset(Vouch_ID_Lag, Vouch_ID_Lag >=0 & Vouch_ID_Lag <=20)
Vouch_ID_Lag <- as.data.frame(sample(Vouch_ID_Lag$Vouch_ID_Lag, 1000))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
#plotting histogram
# ggplot(Vouch_ID_Lag, aes(x = Vouch_ID_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Vouch_ID_Lag)
```
##### 3.4
Then, simulate the lag between IDing and reporting
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

ID_Report_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
ID_Report_Lag <- subset(ID_Report_Lag, ID_Report_Lag >=0 & ID_Report_Lag <=20)
ID_Report_Lag <- as.data.frame(sample(ID_Report_Lag$ID_Report_Lag, 1000))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
#plotting histogram
# ggplot(ID_Report_Lag, aes(x = ID_Report_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(ID_Report_Lag)
```


##### 3.5
Then, simulate the lag between reporting  and when someone makes a resource that is compiled and formatted
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Report_Compile_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
Report_Compile_Lag <- subset(Report_Compile_Lag, Report_Compile_Lag >=0 & Report_Compile_Lag <=20)
Report_Compile_Lag <- as.data.frame(sample(Report_Compile_Lag$Report_Compile_Lag, 1000))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
#plotting histogram
# ggplot(Report_Compile_Lag, aes(x = Report_Compile_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Report_Compile_Lag)
```

So, it may be arbitrary, but the report would compile all species up to that date.

##### 3.6 Okay, lets add up the lags
So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag,

All created by adding or subtracting from the naturalization year
```{r}
sim_spp7$Intro_Year <- sim_spp7$Naturalization_Year - Intro_Nat_Lag$Intro_Nat_Lag
sim_spp7$Detect_Year <- sim_spp7$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag
sim_spp7$Voucher_Year <- sim_spp7$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag
sim_spp7$ID_Year <- sim_spp7$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag
sim_spp7$Report_Year <- sim_spp7$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag


sim_spp7$Compile_Year <- sim_spp7$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag + Report_Compile_Lag$Report_Compile_Lag


sim_spp7$All_Processing_Lags <- Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag
#reordering according to Report Year
sim_spp7 <- sim_spp7 %>% dplyr::select(Species_IDNo, Intro_Year, Naturalization_Year, Detect_Year, Voucher_Year, ID_Year, Report_Year, Compile_Year, All_Processing_Lags)
#sim_spp7
```

### Cropping the Dataset because we can't know about species that haven't been reported yet

What's the lags look like for all species
```{r}
# #plotting histogram
# ggplot(sim_spp7, aes(x = All_Processing_Lags)) +
#   geom_density(fill="#440154FF", color="#e9ecef", alpha=0.6) +
#  # scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
#   theme_classic()
```

MAJOR DIFFERENCE: you get to crop based on the ID year, because you're looking at the vouchers themselves
How many species did we lose that hadn't been reported yet?
```{r}
sim_spp7_crop <- subset(sim_spp7, Compile_Year <= 2020)
```

```{r}
nrow(sim_spp7) -nrow(sim_spp7_crop)
```
```{r}
retrieval_rate = 1
sampled <- as.data.frame(sample(sim_spp7_crop$Species_IDNo, (nrow(sim_spp7_crop)* retrieval_rate)))
colnames(sampled) <- "Species_IDNo"
sim_spp7_crop_sample <- merge(sampled, sim_spp7_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE)
sim_spp7_crop_sample
```

## Step 4. Analyse Trends - A) Voucher Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(sim_spp7_crop_sample$Voucher_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Vouchered = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum)
accumulated_A <- cbind(year_count, accumulated_A)
```

## Step 5. Analyse Trends - B) Report Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp7_crop_sample$Report_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Reported = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_B <- accumulate(year_count$Taxa_Reported, sum)
accumulated_B <- cbind(year_count, accumulated_B)
```
## Step 6. Analyse Trends - C) Intro Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp7_crop_sample$Intro_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Introduced = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_C <- accumulate(year_count$Taxa_Introduced, sum)
accumulated_C <- cbind(year_count, accumulated_C)
```


## Step 6. Analyse Trends - D) Compiled Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp7_crop_sample$Compile_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Compiled = Freq)
year_count

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_D <- accumulate(year_count$Taxa_Compiled, sum)
accumulated_D <- cbind(year_count, accumulated_D)
```

#### Changepoint - Evidence Collection

<!-- ```{r} -->
<!-- plot(accumulated_A$Taxa_Vouchered) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- a1 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- plot(a1, diagnostic = TRUE) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- selectedcp = 4 -->
<!-- #changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)") -->
<!-- changepoint::plot(a1, ncpts = selectedcp) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- changepoint::cpts.full(a1) + (accumulated_A$Year[1] -1) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- firstrow = 5 -->
<!-- secondrow = firstrow +1 -->
<!-- a2 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "Manual",  pen.value=((pen.value.full(a1)[firstrow] + pen.value.full(a1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A)) -->
<!-- chpoints -->
<!-- chpoints + accumulated_A$Year[1] -->
<!-- segmeans <- changepoint::param.est(a2)$mean -->
<!-- segmeans -->
<!-- segvars <- (changepoint::param.est(a2)$variance) -->
<!-- sqrt(segvars) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8]) -->
<!-- seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8]) -->

<!-- segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean) -->
<!-- segwmeans <- segwmeans[!is.na(segwmeans)] -->
<!-- segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar) -->
<!-- segwvars <- segwvars[!is.na(segwvars)] -->
<!-- ``` -->
<!-- ```{r} -->
<!-- startptindex = 3 -->
<!-- endptindex = 4 -->

<!-- print("How many species did we lose that hadn't been reported yet?") -->
<!-- spplostharvest <- nrow(sim_spp7) -nrow(sim_spp7_crop) -->
<!-- spplostharvest -->

<!-- spplostretrieval <- nrow(sim_spp7_crop) - nrow(sim_spp7_crop_sample) -->
<!-- spplostretrieval -->

<!-- print("start seg length, mean, std: ") -->
<!-- chpoints[startptindex] -->
<!-- round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2) -->
<!-- round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2) -->

<!-- print("truncated seg length, mean, std:") -->
<!-- chpoints[endptindex] - chpoints[startptindex] -->
<!-- round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2) -->

<!-- print("ending seg length, mean, std:") -->
<!-- nrow(accumulated_A) - chpoints[endptindex] -->
<!-- round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- changepoint::plot(a2, type='p') -->
<!-- ``` -->



<!-- #### Changepoints - Intro Date -->

<!-- ```{r} -->
<!-- plot(accumulated_C$Taxa_Introduced) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- c1 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- plot(c1, diagnostic = TRUE) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- selectedcp = 4 -->
<!-- #changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)") -->
<!-- changepoint::plot(c1, ncpts = selectedcp) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- changepoint::cpts.full(c1) + as.numeric(accumulated_C$Year[1]) -1 -->
<!-- ``` -->
<!-- ```{r} -->
<!-- firstrow = 3 -->
<!-- secondrow = firstrow +1 -->
<!-- c2 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "Manual",  pen.value=((pen.value.full(c1)[firstrow] + pen.value.full(c1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- chpoints <- c( 0, changepoint::cpts(c2), nrow(accumulated_C)) -->
<!-- chpoints -->
<!-- chpoints + as.numeric(accumulated_C$Year[1]) -->
<!-- segmeans <- changepoint::param.est(c2)$mean -->
<!-- segmeans -->
<!-- segvars <- (changepoint::param.est(c2)$variance) -->
<!-- sqrt(segvars) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8]) -->
<!-- seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8]) -->

<!-- segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean) -->
<!-- segwmeans <- segwmeans[!is.na(segwmeans)] -->
<!-- segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar) -->
<!-- segwvars <- segwvars[!is.na(segwvars)] -->
<!-- ``` -->
<!-- ```{r} -->
<!-- startptindex = 2 -->
<!-- endptindex = 3 -->

<!-- print("How many species did we lose that hadn't been reported yet?") -->
<!-- spplostharvest <- nrow(sim_spp5) -nrow(sim_spp5_crop) -->
<!-- spplostharvest -->

<!-- spplostretrieval <- nrow(sim_spp5_crop) - nrow(sim_spp5_crop_sample) -->
<!-- spplostretrieval -->

<!-- print("start seg length, mean, std: ") -->
<!-- chpoints[startptindex] -->
<!-- round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2) -->
<!-- round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2) -->

<!-- print("truncated seg length, mean, std:") -->
<!-- chpoints[endptindex] - chpoints[startptindex] -->
<!-- round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2) -->

<!-- print("ending seg length, mean, std:") -->
<!-- nrow(accumulated_A) - chpoints[endptindex] -->
<!-- round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- changepoint::plot(c2, type='p') -->
<!-- ``` -->



```{r}
#making into a single dataframe
#accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered)
accumulated_B  <- accumulated_B %>% dplyr::rename(accumulated_species = accumulated_B, No_Taxa = Taxa_Reported)
accumulated_C  <- accumulated_C %>% dplyr::rename(accumulated_species = accumulated_C, No_Taxa = Taxa_Introduced)
accumulated_D  <- accumulated_D %>% dplyr::rename(accumulated_species = accumulated_D, No_Taxa = Taxa_Compiled)
accumulated$Analysis_Type <- "Hypothetical Naturalization Date"
accumulated_A$Analysis_Type <- "Evidence Collection Date"
accumulated_B$Analysis_Type <- "Report Date"
accumulated_C$Analysis_Type <- "Introduction Date"
accumulated_D$Analysis_Type <- "Compiled Date"

accumulated_A_B_C_D <- rbind(accumulated, accumulated_A, accumulated_B, accumulated_C, accumulated_D)
accumulated_A_B_C_D
```

```{r}
# DATA <- accumulated_A_B_C_D %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
# DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
# #plotting total species accumulation over time
# ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(No_Taxa), group = Analysis_Type, color = Analysis_Type)) +
#   geom_smooth( size=1.5, alpha=0.8, level = 0, method = "loess", fullrange=TRUE) +
#   scale_color_viridis(discrete = TRUE) +
#   scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
#   guides(colour = guide_legend(reverse=T)) +
#  theme_classic() +
#   labs(title = "Scenario7 Nat Rate", x = "Year", y = "Number of Species")
# ggsave(filename = "Scenario7_ rate.png", device = "png", scale = 1, width =6, dpi = 600)
```

```{r}
DATA <- accumulated_A_B_C_D# %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
DATA <- subset(DATA, DATA$Analysis_Type != "Report Date")
DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) +
  geom_line( size=1.5, alpha=0.70) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
  guides(colour = guide_legend(reverse=T)) +
 theme_classic() +
  theme(legend.title=element_blank()) +
  labs(title = "Continually Updated Database (with report-compilation lag)", x = "Year", y = "Number of Species")
ggsave(filename = "Scenario7.png", device = "png", scale = 1, width =6, dpi = 600)
```

# Scenario 8: Access Points Introduce different cutoff points: Checklist-style resource compiled 3 years ago.

- prevents leveling off, but excludes species because not all vouchers mentions their status


## Simulate Hypothetical Lag Times

For this scenario, let's say that all lags have a mean of 10 years

```{r}
sim_spp8 <- sim_spp
sim_spp8
```

##### 3.0
Simulate the lag between introduction and naturalization

```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Intro_Nat_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
Intro_Nat_Lag <- subset(Intro_Nat_Lag, Intro_Nat_Lag >=0 & Intro_Nat_Lag <=20)
Intro_Nat_Lag <- as.data.frame(sample(Intro_Nat_Lag$Intro_Nat_Lag, 1000))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"

#plotting histogram
# ggplot(Intro_Nat_Lag, aes(x = Intro_Nat_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Intro_Nat_Lag)
#Nat_Detect_Lag
```

##### 3.1
Then, simulate the lag between naturalization and detection
let's say this takes around 5 years, with some taking a very long time
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Nat_Detect_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
Nat_Detect_Lag <- subset(Nat_Detect_Lag, Nat_Detect_Lag >=0 & Nat_Detect_Lag <=20)
Nat_Detect_Lag <- as.data.frame(sample(Nat_Detect_Lag$Nat_Detect_Lag, 1000))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
#plotting histogram
# ggplot(Nat_Detect_Lag, aes(x = Nat_Detect_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Nat_Detect_Lag)
#Nat_Detect_Lag
```
##### 3.2
Second, simulate the lag between detection and evidence collection
let's say this takes around 1 years, with some taking longer, up to 5 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Detect_Vouch_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Detect_Vouch_Lag) <- "Detect_Vouch_Lag"
Detect_Vouch_Lag <- subset(Detect_Vouch_Lag, Detect_Vouch_Lag >=0 & Detect_Vouch_Lag <=20)
Detect_Vouch_Lag <- as.data.frame(sample(Detect_Vouch_Lag$Detect_Vouch_Lag, 1000))
colnames(Detect_Vouch_Lag) <- "Detect_Vouch_Lag"
#plotting histogram
# ggplot(Detect_Vouch_Lag, aes(x = Detect_Vouch_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Detect_Vouch_Lag)

```

##### 3.3

third, simulate the lag between vouchering and identification
let's say this takes around 1 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Vouch_ID_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
Vouch_ID_Lag <- subset(Vouch_ID_Lag, Vouch_ID_Lag >=0 & Vouch_ID_Lag <=20)
Vouch_ID_Lag <- as.data.frame(sample(Vouch_ID_Lag$Vouch_ID_Lag, 1000))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
#plotting histogram
# ggplot(Vouch_ID_Lag, aes(x = Vouch_ID_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Vouch_ID_Lag)
```
##### 3.4
Then, simulate the lag between IDing and reporting
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

ID_Report_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
ID_Report_Lag <- subset(ID_Report_Lag, ID_Report_Lag >=0 & ID_Report_Lag <=20)
ID_Report_Lag <- as.data.frame(sample(ID_Report_Lag$ID_Report_Lag, 1000))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
#plotting histogram
# ggplot(ID_Report_Lag, aes(x = ID_Report_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(ID_Report_Lag)
```


##### 3.5
Then, simulate the lag between reporting  and when someone makes a resource that is compiled and formatted
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Report_Compile_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
Report_Compile_Lag <- subset(Report_Compile_Lag, Report_Compile_Lag >=0 & Report_Compile_Lag <=20)
Report_Compile_Lag <- as.data.frame(sample(Report_Compile_Lag$Report_Compile_Lag, 1000))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
#plotting histogram
# ggplot(Report_Compile_Lag, aes(x = Report_Compile_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Report_Compile_Lag)
```

So, it may be arbitrary, but the report would compile all species up to that date.

##### 3.6 Okay, lets add up the lags
So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag,

All created by adding or subtracting from the naturalization year
```{r}
sim_spp8$Intro_Year <- sim_spp8$Naturalization_Year - Intro_Nat_Lag$Intro_Nat_Lag
sim_spp8$Detect_Year <- sim_spp8$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag
sim_spp8$Voucher_Year <- sim_spp8$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag
sim_spp8$ID_Year <- sim_spp8$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag
sim_spp8$Report_Year <- sim_spp8$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag


sim_spp8$Compile_Year <- sim_spp8$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag + Report_Compile_Lag$Report_Compile_Lag


sim_spp8$All_Processing_Lags <- Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag
#reordering according to Report Year
sim_spp8 <- sim_spp8 %>% dplyr::select(Species_IDNo, Intro_Year, Naturalization_Year, Detect_Year, Voucher_Year, ID_Year, Report_Year, Compile_Year, All_Processing_Lags)
sim_spp8
```

### Cropping the Dataset because we can't know about species that haven't been reported yet

What's the lags look like for all species
```{r}
# #plotting histogram
# ggplot(sim_spp8, aes(x = All_Processing_Lags)) +
#   geom_density(fill="#440154FF", color="#e9ecef", alpha=0.6) +
#  # scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
#   theme_classic()
```

How many species did we lose that hadn't been reported yet?
```{r}
sim_spp8_crop <- subset(sim_spp8, Report_Year <= 2017)
sim_spp8_crop
```

```{r}
nrow(sim_spp8) -nrow(sim_spp8_crop)
```
```{r}
retrieval_rate = 1
sampled <- as.data.frame(sample(sim_spp8_crop$Species_IDNo, (nrow(sim_spp8_crop)* retrieval_rate)))
colnames(sampled) <- "Species_IDNo"
sim_spp8_crop_sample <- merge(sampled, sim_spp8_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE)
sim_spp8_crop_sample
```


## Step 4. Analyse Trends - A) Voucher Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(sim_spp8_crop_sample$Voucher_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Vouchered = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum)
accumulated_A <- cbind(year_count, accumulated_A)
accumulated_A
```

## Step 5. Analyse Trends - B) Report Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp8_crop_sample$Report_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Reported = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_B <- accumulate(year_count$Taxa_Reported, sum)
accumulated_B <- cbind(year_count, accumulated_B)
```
## Step 6. Analyse Trends - C) Intro Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp8_crop_sample$Intro_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Introduced = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_C <- accumulate(year_count$Taxa_Introduced, sum)
accumulated_C <- cbind(year_count, accumulated_C)
```


## Step 6. Analyse Trends - D) Compiled Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp8_crop_sample$Compile_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Compiled = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_D <- accumulate(year_count$Taxa_Compiled, sum)
accumulated_D <- cbind(year_count, accumulated_D)
```

#### Changepoint - Evidence Collection


<!-- ```{r} -->
<!-- plot(accumulated_A$Taxa_Vouchered) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- a1 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- plot(a1, diagnostic = TRUE) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- selectedcp = 4 -->
<!-- #changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)") -->
<!-- changepoint::plot(a1, ncpts = selectedcp) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- changepoint::cpts.full(a1) + (accumulated_A$Year[1] -1) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- firstrow = 4 -->
<!-- secondrow = firstrow +1 -->
<!-- a2 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "Manual",  pen.value=((pen.value.full(a1)[firstrow] + pen.value.full(a1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A)) -->
<!-- chpoints -->
<!-- chpoints + accumulated_A$Year[1] -->
<!-- segmeans <- changepoint::param.est(a2)$mean -->
<!-- segmeans -->
<!-- segvars <- (changepoint::param.est(a2)$variance) -->
<!-- sqrt(segvars) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8]) -->
<!-- seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8]) -->

<!-- segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean) -->
<!-- segwmeans <- segwmeans[!is.na(segwmeans)] -->
<!-- segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar) -->
<!-- segwvars <- segwvars[!is.na(segwvars)] -->
<!-- ``` -->
<!-- ```{r} -->
<!-- startptindex = 3 -->
<!-- endptindex = 4 -->

<!-- print("How many species did we lose that hadn't been reported yet?") -->
<!-- spplostharvest <- nrow(sim_spp8) -nrow(sim_spp8_crop) -->
<!-- spplostharvest -->

<!-- spplostretrieval <- nrow(sim_spp8_crop) - nrow(sim_spp8_crop_sample) -->
<!-- spplostretrieval -->

<!-- print("start seg length, mean, std: ") -->
<!-- chpoints[startptindex] -->
<!-- round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2) -->
<!-- round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2) -->

<!-- print("truncated seg length, mean, std:") -->
<!-- chpoints[endptindex] - chpoints[startptindex] -->
<!-- round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2) -->

<!-- print("ending seg length, mean, std:") -->
<!-- nrow(accumulated_A) - chpoints[endptindex] -->
<!-- round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- changepoint::plot(a2, type='p') -->
<!-- ``` -->



<!-- #### Changepoints - Intro Date -->

<!-- ```{r} -->
<!-- plot(accumulated_C$Taxa_Introduced) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- c1 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- plot(c1, diagnostic = TRUE) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- selectedcp = 4 -->
<!-- #changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)") -->
<!-- changepoint::plot(c1, ncpts = selectedcp) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- changepoint::cpts.full(c1) + as.numeric(accumulated_C$Year[1]) -1 -->
<!-- ``` -->
<!-- ```{r} -->
<!-- firstrow = 3 -->
<!-- secondrow = firstrow +1 -->
<!-- c2 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "Manual",  pen.value=((pen.value.full(c1)[firstrow] + pen.value.full(c1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- chpoints <- c( 0, changepoint::cpts(c2), nrow(accumulated_C)) -->
<!-- chpoints -->
<!-- chpoints + as.numeric(accumulated_C$Year[1]) -->
<!-- segmeans <- changepoint::param.est(c2)$mean -->
<!-- segmeans -->
<!-- segvars <- (changepoint::param.est(c2)$variance) -->
<!-- sqrt(segvars) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8]) -->
<!-- seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8]) -->

<!-- segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean) -->
<!-- segwmeans <- segwmeans[!is.na(segwmeans)] -->
<!-- segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar) -->
<!-- segwvars <- segwvars[!is.na(segwvars)] -->
<!-- ``` -->
<!-- ```{r} -->
<!-- startptindex = 2 -->
<!-- endptindex = 3 -->

<!-- print("How many species did we lose that hadn't been reported yet?") -->
<!-- spplostharvest <- nrow(sim_spp8) -nrow(sim_spp8_crop) -->
<!-- spplostharvest -->

<!-- spplostretrieval <- nrow(sim_spp8_crop) - nrow(sim_spp8_crop_sample) -->
<!-- spplostretrieval -->

<!-- print("start seg length, mean, std: ") -->
<!-- chpoints[startptindex] -->
<!-- round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2) -->
<!-- round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2) -->

<!-- print("truncated seg length, mean, std:") -->
<!-- chpoints[endptindex] - chpoints[startptindex] -->
<!-- round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2) -->

<!-- print("ending seg length, mean, std:") -->
<!-- nrow(accumulated_A) - chpoints[endptindex] -->
<!-- round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- changepoint::plot(c2, type='p') -->
<!-- ``` -->



```{r}
#making into a single dataframe
#accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered)
accumulated_B  <- accumulated_B %>% dplyr::rename(accumulated_species = accumulated_B, No_Taxa = Taxa_Reported)
accumulated_C  <- accumulated_C %>% dplyr::rename(accumulated_species = accumulated_C, No_Taxa = Taxa_Introduced)
accumulated_D  <- accumulated_D %>% dplyr::rename(accumulated_species = accumulated_D, No_Taxa = Taxa_Compiled)
accumulated$Analysis_Type <- "Hypothetical Naturalization Date"
accumulated_A$Analysis_Type <- "Evidence Collection Date"
accumulated_B$Analysis_Type <- "Report Date"
accumulated_C$Analysis_Type <- "Introduction Date"
accumulated_D$Analysis_Type <- "Compiled Date"

accumulated_A_B_C_D <- rbind(accumulated, accumulated_A, accumulated_B, accumulated_C, accumulated_D)
accumulated_A_B_C_D
```

```{r}
# DATA <- accumulated_A_B_C_D %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
# DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
# #plotting total species accumulation over time
# ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(No_Taxa), group = Analysis_Type, color = Analysis_Type)) +
#   geom_smooth( size=1.5, alpha=0.8, level = 0, method = "loess", fullrange=TRUE) +
#   scale_color_viridis(discrete = TRUE) +
#   scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
#   guides(colour = guide_legend(reverse=T)) +
#  theme_classic() +
#   labs(title = "Scenario8 Nat Rate", x = "Year", y = "Number of Species")
# ggsave(filename = "Scenario8_ rate.png", device = "png", scale = 1, width =6, dpi = 600)
```

```{r}
DATA <- accumulated_A_B_C_D# %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
DATA <- subset(DATA, DATA$Analysis_Type != "Report Date")
DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) +
  geom_line( size=1.5, alpha=0.70) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
  guides(colour = guide_legend(reverse=T)) +
 theme_classic() +
  theme(legend.title=element_blank()) +
  labs(title = "Compiled Resource published in 2017", x = "Year", y = "Number of Species")
ggsave(filename = "Scenario8.png", device = "png", scale = 1, width =6, dpi = 600)
```

# Scenario 9: Access Points Introduce different cutoff points: Checklist-style resource compiled 30 years ago.

- prevents leveling off, but excludes species because not all vouchers mentions their status


## Simulate Hypothetical Lag Times

For this scenario, let's say that all lags have a mean of 10 years

```{r}
sim_spp9 <- sim_spp
sim_spp9
```

##### 3.0
Simulate the lag between introduction and naturalization

```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Intro_Nat_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"
Intro_Nat_Lag <- subset(Intro_Nat_Lag, Intro_Nat_Lag >=0 & Intro_Nat_Lag <=20)
Intro_Nat_Lag <- as.data.frame(sample(Intro_Nat_Lag$Intro_Nat_Lag, 1000))
colnames(Intro_Nat_Lag) <- "Intro_Nat_Lag"

#plotting histogram
# ggplot(Intro_Nat_Lag, aes(x = Intro_Nat_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Intro_Nat_Lag)
#Nat_Detect_Lag
```

##### 3.1
Then, simulate the lag between naturalization and detection
let's say this takes around 5 years, with some taking a very long time
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Nat_Detect_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
Nat_Detect_Lag <- subset(Nat_Detect_Lag, Nat_Detect_Lag >=0 & Nat_Detect_Lag <=20)
Nat_Detect_Lag <- as.data.frame(sample(Nat_Detect_Lag$Nat_Detect_Lag, 1000))
colnames(Nat_Detect_Lag) <- "Nat_Detect_Lag"
#plotting histogram
# ggplot(Nat_Detect_Lag, aes(x = Nat_Detect_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Nat_Detect_Lag)
#Nat_Detect_Lag
```
##### 3.2
Second, simulate the lag between detection and evidence collection
let's say this takes around 1 years, with some taking longer, up to 5 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Detect_Vouch_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Detect_Vouch_Lag) <- "Detect_Vouch_Lag"
Detect_Vouch_Lag <- subset(Detect_Vouch_Lag, Detect_Vouch_Lag >=0 & Detect_Vouch_Lag <=20)
Detect_Vouch_Lag <- as.data.frame(sample(Detect_Vouch_Lag$Detect_Vouch_Lag, 1000))
colnames(Detect_Vouch_Lag) <- "Detect_Vouch_Lag"
#plotting histogram
# ggplot(Detect_Vouch_Lag, aes(x = Detect_Vouch_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Detect_Vouch_Lag)

```

##### 3.3

third, simulate the lag between vouchering and identification
let's say this takes around 1 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Vouch_ID_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
Vouch_ID_Lag <- subset(Vouch_ID_Lag, Vouch_ID_Lag >=0 & Vouch_ID_Lag <=20)
Vouch_ID_Lag <- as.data.frame(sample(Vouch_ID_Lag$Vouch_ID_Lag, 1000))
colnames(Vouch_ID_Lag) <- "Vouch_ID_Lag"
#plotting histogram
# ggplot(Vouch_ID_Lag, aes(x = Vouch_ID_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Vouch_ID_Lag)
```
##### 3.4
Then, simulate the lag between IDing and reporting
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

ID_Report_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
ID_Report_Lag <- subset(ID_Report_Lag, ID_Report_Lag >=0 & ID_Report_Lag <=20)
ID_Report_Lag <- as.data.frame(sample(ID_Report_Lag$ID_Report_Lag, 1000))
colnames(ID_Report_Lag) <- "ID_Report_Lag"
#plotting histogram
# ggplot(ID_Report_Lag, aes(x = ID_Report_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(ID_Report_Lag)
```


##### 3.5
Then, simulate the lag between reporting  and when someone makes a resource that is compiled and formatted
let's say this takes around 2 years, with some taking longer, up to 10 years
```{r}
mean = 10
SD = 3.5
skew = 0
params <- sn::cp2dp(c(mean, SD, skew), "SN")

Report_Compile_Lag <- as.data.frame(round(sn::rsn(10000, dp =params), digits = 0))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
Report_Compile_Lag <- subset(Report_Compile_Lag, Report_Compile_Lag >=0 & Report_Compile_Lag <=20)
Report_Compile_Lag <- as.data.frame(sample(Report_Compile_Lag$Report_Compile_Lag, 1000))
colnames(Report_Compile_Lag) <- "Report_Compile_Lag"
#plotting histogram
# ggplot(Report_Compile_Lag, aes(x = Report_Compile_Lag)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# summary(Report_Compile_Lag)
```


##### 3.6 Okay, lets add up the lags
So, the timing of the collected voucher AND final "new record report" is just the year the species naturalized + lag,

All created by adding or subtracting from the naturalization year
```{r}
sim_spp9$Intro_Year <- sim_spp9$Naturalization_Year - Intro_Nat_Lag$Intro_Nat_Lag
sim_spp9$Detect_Year <- sim_spp9$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag
sim_spp9$Voucher_Year <- sim_spp9$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag
sim_spp9$ID_Year <- sim_spp9$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag
sim_spp9$Report_Year <- sim_spp9$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag


sim_spp9$Compile_Year <- sim_spp9$Naturalization_Year + Nat_Detect_Lag$Nat_Detect_Lag + Detect_Vouch_Lag$Detect_Vouch_Lag + Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag + Report_Compile_Lag$Report_Compile_Lag


sim_spp9$All_Processing_Lags <- Vouch_ID_Lag$Vouch_ID_Lag + ID_Report_Lag$ID_Report_Lag
#reordering according to Report Year
sim_spp9 <- sim_spp9 %>% dplyr::select(Species_IDNo, Intro_Year, Naturalization_Year, Detect_Year, Voucher_Year, ID_Year, Report_Year, Compile_Year, All_Processing_Lags)
sim_spp9
```

### Cropping the Dataset because we can't know about species that haven't been reported yet

What's the lags look like for all species
```{r}
# #plotting histogram
# ggplot(sim_spp9, aes(x = All_Processing_Lags)) +
#   geom_density(fill="#440154FF", color="#e9ecef", alpha=0.6) +
#  # scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
#   theme_classic()
```

MAJOR DIFFERENCE: you get to crop based on the ID year, because you're looking at the vouchers themselves
How many species did we lose that hadn't been reported yet?
```{r}
sim_spp9_crop <- subset(sim_spp9, Report_Year <= 1995)
sim_spp9_crop
```

```{r}
nrow(sim_spp9) -nrow(sim_spp9_crop)
```
```{r}
retrieval_rate = 1
sampled <- as.data.frame(sample(sim_spp9_crop$Species_IDNo, (nrow(sim_spp9_crop)* retrieval_rate)))
colnames(sampled) <- "Species_IDNo"
sim_spp9_crop_sample <- merge(sampled, sim_spp9_crop, by= "Species_IDNo", all.x = TRUE, all.y = FALSE)
sim_spp9_crop_sample
```

Can we find the best model to fit this line?  Look at Seebens' code.

## Step 4. Analyse Trends - A) Voucher Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they we vouchered
year_count <- as.data.frame(table(sim_spp9_crop_sample$Voucher_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Vouchered = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0

#calculating species accumulation over time
accumulated_A <- accumulate(year_count$Taxa_Vouchered, sum)
accumulated_A <- cbind(year_count, accumulated_A)
```

## Step 5. Analyse Trends - B) Report Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp9_crop_sample$Report_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Reported = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_B <- accumulate(year_count$Taxa_Reported, sum)
accumulated_B <- cbind(year_count, accumulated_B)
```
## Step 6. Analyse Trends - C) Intro Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp9_crop_sample$Intro_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Introduced = Freq)


#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_C <- accumulate(year_count$Taxa_Introduced, sum)
accumulated_C <- cbind(year_count, accumulated_C)

```


## Step 6. Analyse Trends - D) Compiled Year
```{r}
#just making a vector of years from 1920 to 2020
alltheyears <- as.data.frame(seq(from = 1920, to = 2019))
colnames(alltheyears) <- "Year"

#grouping my simulated species by the year they were reported
year_count <- as.data.frame(table(sim_spp9_crop_sample$Compile_Year)) %>%
# and renaming columns
  dplyr::rename(Year = Var1, Taxa_Compiled = Freq)

#merging with all the years
year_count <-merge(alltheyears, year_count, all = TRUE)
#making the NA's become 0s, because 0 new species were vouchered in those years
year_count[is.na(year_count)] = 0
#calculating species accumulation over time
accumulated_D <- accumulate(year_count$Taxa_Compiled, sum)
accumulated_D <- cbind(year_count, accumulated_D)
```

#### Changepoint - Evidence Collection


<!-- ```{r} -->
<!-- plot(accumulated_A$Taxa_Vouchered) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- a1 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- plot(a1, diagnostic = TRUE) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- selectedcp = 4 -->
<!-- #changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)") -->
<!-- changepoint::plot(a1, ncpts = selectedcp) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- changepoint::cpts.full(a1) + (accumulated_A$Year[1] -1) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- firstrow = 4 -->
<!-- secondrow = firstrow +1 -->
<!-- a2 <- changepoint::cpt.meanvar(accumulated_A$Taxa_Vouchered, penalty = "Manual",  pen.value=((pen.value.full(a1)[firstrow] + pen.value.full(a1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- chpoints <- c( 0, changepoint::cpts(a2), nrow(accumulated_A)) -->
<!-- chpoints -->
<!-- chpoints + accumulated_A$Year[1] -->
<!-- segmeans <- changepoint::param.est(a2)$mean -->
<!-- segmeans -->
<!-- segvars <- (changepoint::param.est(a2)$variance) -->
<!-- sqrt(segvars) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8]) -->
<!-- seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8]) -->

<!-- segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean) -->
<!-- segwmeans <- segwmeans[!is.na(segwmeans)] -->
<!-- segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar) -->
<!-- segwvars <- segwvars[!is.na(segwvars)] -->
<!-- ``` -->
<!-- ```{r} -->
<!-- startptindex = 3 -->
<!-- endptindex = 4 -->

<!-- print("How many species did we lose that hadn't been reported yet?") -->
<!-- spplostharvest <- nrow(sim_spp8) -nrow(sim_spp8_crop) -->
<!-- spplostharvest -->

<!-- spplostretrieval <- nrow(sim_spp8_crop) - nrow(sim_spp8_crop_sample) -->
<!-- spplostretrieval -->

<!-- print("start seg length, mean, std: ") -->
<!-- chpoints[startptindex] -->
<!-- round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2) -->
<!-- round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2) -->

<!-- print("truncated seg length, mean, std:") -->
<!-- chpoints[endptindex] - chpoints[startptindex] -->
<!-- round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2) -->

<!-- print("ending seg length, mean, std:") -->
<!-- nrow(accumulated_A) - chpoints[endptindex] -->
<!-- round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- changepoint::plot(a2, type='p') -->
<!-- ``` -->



<!-- #### Changepoints - Intro Date -->

<!-- ```{r} -->
<!-- plot(accumulated_C$Taxa_Introduced) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- c1 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "CROPS",  pen.value=c(5,500), method = "PELT", Q = 25, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- plot(c1, diagnostic = TRUE) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- selectedcp = 4 -->
<!-- #changepoint::plot(a1, cpt.width = 4, col = "black", cpt.col = "blue", ylab = "# of Taxa Vouchered", xlab = "Years (since beginning of time window)") -->
<!-- changepoint::plot(c1, ncpts = selectedcp) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- changepoint::cpts.full(c1) + as.numeric(accumulated_C$Year[1]) -1 -->
<!-- ``` -->
<!-- ```{r} -->
<!-- firstrow = 3 -->
<!-- secondrow = firstrow +1 -->
<!-- c2 <- changepoint::cpt.meanvar(accumulated_C$Taxa_Introduced, penalty = "Manual",  pen.value=((pen.value.full(c1)[firstrow] + pen.value.full(c1)[secondrow]) /2) , method = "PELT", Q = 10, test.stat = "Normal", class = TRUE, param.estimates = TRUE, minseglen = 5) -->

<!-- chpoints <- c( 0, changepoint::cpts(c2), nrow(accumulated_C)) -->
<!-- chpoints -->
<!-- chpoints + as.numeric(accumulated_C$Year[1]) -->
<!-- segmeans <- changepoint::param.est(c2)$mean -->
<!-- segmeans -->
<!-- segvars <- (changepoint::param.est(c2)$variance) -->
<!-- sqrt(segvars) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- seg1wmean <- segmeans[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wmean <- segmeans[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wmean <- segmeans[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wmean <- segmeans[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wmean <- segmeans[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wmean <- segmeans[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wmean <- segmeans[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg8wmean <- segmeans[8]* (chpoints[9] - chpoints[8]) -->
<!-- seg1wvar <- segvars[1]* (chpoints[2]- chpoints[1]) -->
<!-- seg2wvar <- segvars[2]* (chpoints[3] - chpoints[2]) -->
<!-- seg3wvar <- segvars[3]* (chpoints[4] - chpoints[3]) -->
<!-- seg4wvar <- segvars[4]* (chpoints[5] - chpoints[4]) -->
<!-- seg5wvar <- segvars[5]* (chpoints[6] - chpoints[5]) -->
<!-- seg6wvar <- segvars[6]* (chpoints[7] - chpoints[6]) -->
<!-- seg7wvar <- segvars[7]* (chpoints[8] - chpoints[7]) -->
<!-- seg7wvar <- segvars[8]* (chpoints[9] - chpoints[8]) -->

<!-- segwmeans <- c(seg1wmean, seg2wmean, seg3wmean, seg4wmean, seg5wmean, seg6wmean, seg7wmean) -->
<!-- segwmeans <- segwmeans[!is.na(segwmeans)] -->
<!-- segwvars <- c(seg1wvar, seg2wvar, seg3wvar, seg4wvar, seg5wvar, seg6wvar, seg7wvar) -->
<!-- segwvars <- segwvars[!is.na(segwvars)] -->
<!-- ``` -->
<!-- ```{r} -->
<!-- startptindex = 2 -->
<!-- endptindex = 3 -->

<!-- print("How many species did we lose that hadn't been reported yet?") -->
<!-- spplostharvest <- nrow(sim_spp8) -nrow(sim_spp8_crop) -->
<!-- spplostharvest -->

<!-- spplostretrieval <- nrow(sim_spp9_crop) - nrow(sim_spp9_crop_sample) -->
<!-- spplostretrieval -->

<!-- print("start seg length, mean, std: ") -->
<!-- chpoints[startptindex] -->
<!-- round(sum(segwmeans[1:(startptindex-1)]) / chpoints[startptindex], 2) -->
<!-- round(sqrt(sum(segwvars[1:(startptindex-1)]) / chpoints[startptindex]), 2) -->

<!-- print("truncated seg length, mean, std:") -->
<!-- chpoints[endptindex] - chpoints[startptindex] -->
<!-- round(sum(segwmeans[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[startptindex:(endptindex -1)]) / (chpoints[endptindex] - chpoints[startptindex])), 2) -->

<!-- print("ending seg length, mean, std:") -->
<!-- nrow(accumulated_A) - chpoints[endptindex] -->
<!-- round(sum(segwmeans[endptindex:length(segwmeans)]) / (chpoints[length(chpoints)]- chpoints[endptindex]), 2) -->
<!-- round(sqrt(sum(segwvars[endptindex:length(segwvars)]) / (chpoints[length(chpoints)]- chpoints[endptindex])), 2) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- changepoint::plot(c2, type='p') -->
<!-- ``` -->



```{r}
#making into a single dataframe
#accumulated  <- accumulated %>% dplyr::rename(accumulated_species = accumulated, No_Taxa = Taxa_Naturalized)
accumulated_A  <- accumulated_A %>% dplyr::rename(accumulated_species = accumulated_A, No_Taxa = Taxa_Vouchered)
accumulated_B  <- accumulated_B %>% dplyr::rename(accumulated_species = accumulated_B, No_Taxa = Taxa_Reported)
accumulated_C  <- accumulated_C %>% dplyr::rename(accumulated_species = accumulated_C, No_Taxa = Taxa_Introduced)
accumulated_D  <- accumulated_D %>% dplyr::rename(accumulated_species = accumulated_D, No_Taxa = Taxa_Compiled)
accumulated$Analysis_Type <- "Hypothetical Naturalization Date"
accumulated_A$Analysis_Type <- "Evidence Collection Date"
accumulated_B$Analysis_Type <- "Report Date"
accumulated_C$Analysis_Type <- "Introduction Date"
accumulated_D$Analysis_Type <- "Compiled Date"

accumulated_A_B_C_D <- rbind(accumulated, accumulated_A, accumulated_B, accumulated_C, accumulated_D)
accumulated_A_B_C_D
```

```{r}
# DATA <- accumulated_A_B_C_D %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
# DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
# #plotting total species accumulation over time
# ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(No_Taxa), group = Analysis_Type, color = Analysis_Type)) +
#   geom_smooth( size=1.5, alpha=0.8, level = 0, method = "loess", fullrange=TRUE) +
#   scale_color_viridis(discrete = TRUE) +
#   scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
#   guides(colour = guide_legend(reverse=T)) +
#  theme_classic() +
#   labs(title = "Scenario9 Nat Rate", x = "Year", y = "Number of Species")
# ggsave(filename = "Scenario9_ rate.png", device = "png", scale = 1, width =6, dpi = 600)
```

```{r}
DATA <- accumulated_A_B_C_D# %>% mutate(Analysis_Type = fct_reorder(Analysis_Type, accumulated_species))
DATA <- subset(DATA, DATA$Analysis_Type != "Report Date")
DATA <- subset(DATA, DATA$Analysis_Type != "Compiled Date")
#plotting total species accumulation over time
ggplot(DATA, aes(x=as.numeric(Year), y=as.numeric(accumulated_species), group = Analysis_Type, color = Analysis_Type)) +
  geom_line( size=1.5, alpha=0.70) +
  scale_color_viridis(discrete = TRUE) +
  scale_x_continuous(breaks = c(1920, 1940, 1960, 1980, 2000, 2020)) +
  guides(colour = guide_legend(reverse=T)) +
 theme_classic() +
  theme(legend.title=element_blank()) +
  labs(title = "Access Point = Compiled Resource published in 1995", x = "Year", y = "Number of Species")
ggsave(filename = "Scenario9.png", device = "png", scale = 1, width =6, dpi = 600)
```

